### 2024-08-19 raster.qmd
First commit

---
execute:
  warning: false
  message: false
---

# National WUI map

```{r}
#| label: libraries

# Libraries
library(tidyverse)
library(terra)
library(tidyterra)
library(sf)
library(patchwork)
library(gt)
library(gtExtras)
library(purrr)

terraOptions(progress = 0)

```

## Functions

In addition to the packages, we define our own functions for the analysis:

```{r}
#| label: aux-functions

# Function: get_perc_veg_buffer ----
#' Function for calculating if the percentage of vegetation within a buffer area is higher than a threshold 
#' @description It returns a raster indicating if the percentage of flammable vegetation within a circumference, from the centre of the grid (buffer), higher than a threshold (YES = 1, NO = 0)
#' @param .veg raster layer indicating if the vegetation is considered as fuel or not (YES = 1, NO = 0). 
#' @param .radio_buffer Radio of the circumference.
#' @param .threshold Proportion: from 0-1 (default  = 0.5; i.e., 50%) 
#' @param .cache_path File path to save the results 
get_perc_veg_buffer <- function(.veg,
                                .radio_buffer,
                                .threshold = 0.5){

  # Percentage-vegetation 
  fwm <- terra::focalMat(.veg, d = radio_buffer, type = "circle", fillNA = TRUE)
  focal_perc <- terra::focal(.veg, w = fwm, fun = "sum", na.rm = TRUE)

  # Higher than threshold (YES = 1, No = 0)
  wv_perc <- focal_perc
  wv_perc[wv_perc <  .threshold] <- 0
  wv_perc[wv_perc >= .threshold] <- 1

  return(wv_perc)

}

# Function: get_large_polygons ----
#' Function for removing small areas from a raster file. Used for calculating large forest areas (e.g., > 5 km2).  
#' @description It returns a raster with all the polygons larger than a minimal area (YES = 1, NO = 0)
#' @param .x starts object (i.e., 1 - fuel vegetation, 0 - no fuel vegetation). 
#' @param .min_area Minimum area [in km2] for considering an vegetated area as large  (Default 5 km^2)
#' @param .resolution Resolution of the final raster (default - 100 x 100 m)
get_large_polygons <- function(.x,               
                               .min_area = 5,    
                               .resolution = 100) {

  df_v <- .x |> 
    # Convert to polygons
    st_as_sf(as_points = FALSE, merge = TRUE) %>% 
    # Calculate the area of the polygons
    mutate(area = units::set_units(st_area(.), "km2")) |> 
    # Remove lower than 5 km2
    filter(area > units::set_units(.min_area, "km2")) |> 
    # Set columns names
    setNames(c("large_wild_veg", "geometry", "area")) |> 
    # Remove area column
    select(-area) |> 
    # transform to SpatVector  
    vect() |> 
    # aggregate in one polygon
    aggregate(by = "large_wild_veg") |> 
    select(large_wild_veg)

  # Rasterize
  df_geom <- rast(df_v, resolution = .resolution)
  df_r <- rasterize(df_v, df_geom)

  return(df_r)

}

# get_dist_max ----
#' Get cells within a distance to vegetated area (YES --> 1; NO --> 0)  
#' @description It returns a raster with indicating the cells that are within the distance to the vegetated areas (YES = 1, NO = 0)
#' @param .x Raster with distances (from `terra::distance`) 
#' @param .max_distance Maximun distance in m (e.g., 2400 m)
get_dist_max <- function(.x, .max_dist) {

  dist <- .x
  dist[dist <= .max_dist] <- 1
  dist[dist >  .max_dist] <- 0

  return(dist)

}

# Function: get_wui_raster ----
#' Function for WUI classification
#' @description it returns a raster with WUI areas (Interface or Intermix)
#' @param .veg_perc raster layer indicating if the percentage of vegetation in
#' the buffer area is higher than a threshold (e.g., > 50%). (YES = 1, NO = 0). 
#' @param .veg_dist raster layer indicating if the a grid cell is within a 
#' distance (e.g., 2.4 km) of a large area of flammable vegetation (YES = 1, NO = 0)
#' @param .pot_wui raster layer with potential WUI grids based on building density (`wui_potential`).
#' @param .region SpatVector layer with the region area (`region_v`) 
get_wui_raster <- function(.veg_perc,
                           .veg_dist,
                           .pot_wui = wui_potential,
                           .region = region_v){

  # Potential WUI grid ----
  wui_potential <- .pot_wui
  wui_potential <- as.factor(wui_potential)

  # Match extents
  perc <- resample(.veg_perc, wui_potential, method = "mode")
  dist <- resample(.veg_dist, wui_potential, method = "mode")

  # Intermix 
  wui <- perc + wui_potential # 2 = WUI intermix
  wui[wui == 1] <- 0          # 1 = Potential interface --> 0

  # Interface 
  wui <- wui + dist
  wui[wui == 0] <- NA # No WUI
  wui[wui == 1] <- 1  # Interface
  wui[wui == 2] <- 2  # Intermix
  wui[wui == 3] <- 2  # Intermix

  # NAs as 0 inside the study area (otherwise NA)
  wui <- ifel(is.na(wui), 0, wui)

  # Mask to region (e.g., NA = ourside norway)
  wui <- terra::mask(wui, .region)

  # Identify levels (1 = Interface, 2 = Intermix)
  levels(wui) <- data.frame(id = 0:2,
                            wui_class = c("No_wui", "Interface", "Intermix")
                            )

  # Add colour table 
  coltab(wui) <- data.frame(id = 0:2, 
                            col = c("#E7EEE7", "#FF6633", "#003366"))

  # Return SpatRaster
  set.values(wui)
  return(wui)

}

# add_colours_wui 
#' Function for adding WUI colours in a SpatRaster 
#' @description Add colours to a WUI raster layer (No_WUI, Interface, Intermix)
#' @param .veg_perc raster layer indicating if the percentage of vegetation in
add_colours_wui <- function(.x) {

  r <- .x
  # Identify levels (1 = Interface, 2 = Intermix)
  df_lev <- data.frame(id = 0:2, wui_class = c("No_wui", "Interface", "Intermix"))
  df_lev_lst <- replicate(nlyr(r), df_lev, simplify = FALSE)
  levels(r) <- df_lev_lst
  # Add colour table 
  df_col <- data.frame(id = 0:2, col = c("#E7EEE7", "#FF6633", "#003366"))
  df_col_lst <- replicate(nlyr(r), df_col, simplify = FALSE)
  coltab(r) <- df_col_lst

  return(r)

}

```

## Data preprocessing

### Parameter

```{r}
#| label: Parameters-definition

raster_resolution <- 250 # m
radio_buffer <- 500 # m
area_buffer <- pi * (radio_buffer/1000)^2 # km2

```

At national scale, we have performed a raster analysis by grids cells of `r raster_resolution` m x `r raster_resolution` m.

### Study area

We load the last release of the county data from GISCO (`year = 2021`), with a resolution of `1:1 million` and `crs= 3035` (@fig-region-raster). Due to size of the file, we perform the analysis by tiles.

```{r}
#| label: fig-region-raster
#| fig-cap: Study area (Norway and Sweden)

# Create local directory for caching big files
options(gisco_cache_dir = "C:/GISCO_spatial_data")

# Counties ----
country_list <- c("NO", "SE")
counties <- giscoR::gisco_get_nuts(country = country_list,
                                   year = "2021",
                                   nuts_level = 3,
                                   epsg = "3035",
                                   resolution = "01") |> 
  filter(NUTS_NAME != "Svalbard") |> 
  filter(NUTS_NAME != "Jan Mayen") |> 
  select(NUTS_ID, NUTS_NAME, CNTR_CODE)

# Study region (SF object) ----
region_sf <- counties |> 
  summarize()

country_sf <- counties |>
  group_by(CNTR_CODE) |> 
  summarize()

# # Rogaland ----
# rogaland <- counties |> 
#   filter(NUTS_NAME == "Rogaland")

# Transform to terra::SpatVector (region_v) and SpatRaster (region_r)) -----
region_v <- vect(region_sf)
geometry_raster <- rast(region_v, resolution = raster_resolution)
region_r <- rasterize(region_v, geometry_raster)

no_v <- country_sf |> 
  filter(CNTR_CODE == "NO") |> 
  vect()

se_v <- country_sf |> 
  filter(CNTR_CODE == "SE") |> 
  vect()

# Tiles 
raster_tiles <- rast(nrows = 15,
                     ncols = 15,
                     ext(region_r),
                     crs = "EPSG:3035")

dir.create("data/processed_data/tiles")
dir.create("data/processed_data/tiles/region")

region_tiles <- makeTiles(
  x = region_r,
  y = raster_tiles,
  filename = "data/processed_data/tiles/region/region_r_.tif",
  na.rm = TRUE,
  extend = TRUE
)

region_vrt <- vrt(region_tiles)

# Plot Study area ----

# world 
world <- rworldmap::getMap(resolution = "high") |> 
  st_as_sf() |> 
  st_transform(crs = 3035)

p1 <- ggplot() +
  geom_sf(data = world) +
  geom_sf(data = country_sf, fill = c("#D55E00", "#4E84C4")) +
  coord_sf(xlim = c(2500000, 5402285),
           ylim = c(1250000, 5500000),
           expand = FALSE) +
  theme_bw() +
  theme(panel.background = element_rect(fill = "#E7EEE7")) +
  theme(panel.grid = element_blank(),
        axis.text  = element_blank(),
        axis.ticks = element_blank())

p2 <- ggplot() +
  geom_sf(data = region_sf, fill = "#52854C") +
  geom_sf(data = country_sf, fill = NA) +
  theme_bw()

p2 + inset_element(p1, 0.03, 0.70, 0.6, 0.95, align_to = 'full')

```

### Building density

We have downloaded the building data from OpenStreetMaps (geofabrik: [link](https://download.geofabrik.de/europe/)), in pbf format. The version of the datasets are:

+-------------------------------------------------------------------------------------+-------------------+-----------+
| **Name**                                                                            | **Last modified** | **Size**  |
+-------------------------------------------------------------------------------------+-------------------+-----------+
| [norway-latest-osm.pbf](https://download.geofabrik.de/europe/norway-latest.osm.pbf) | 2023-08-17 04:37  | 1.2G      |
+-------------------------------------------------------------------------------------+-------------------+-----------+
| [sweden-latest-osm.pbf](https://download.geofabrik.de/europe/sweden-latest.osm.pbf) | 2023-11-16 04:50  | 656M      |
+-------------------------------------------------------------------------------------+-------------------+-----------+

They are big files and we have not put them on GitHub. If you wanted to replicate the document, you would need to download the file and save it with the same path as here, or change the path in the code (NOTE: you may have small differences in the results if OSM updates the data). To calculate the building density we do the following steps:

-   Transform the `.pbf` file to `.gpkg`, extracting only `multipolygons` (run only one time)
-   Read building polygons
-   Calculate the centroid of each building, assigning a value of 1 to the building (`building = 1`)
-   Calculate the building density in a grid cell of `r raster_resolution` m x `r raster_resolution` m (it is done by tiles). It is the `sum` of all the buildings (`building = 1`) within a grid cell
-   Merge all tiles in one raster (`terra::mosaic`)
-   Smooth the data with a circular moving window of `r radio_buffer` m; i.e., buffer area (`terra::focal`)
-   The density building of a grid cell is therefore the number of buildings in the buffer area divided by the area of the buffer. Potential WUI areas are those grid cells with a building density higher than 6.17 #/km\^2

```{r}
#| label: building-density
#| cache: true
#| warning: false
#| message: false

# # Transform OSM data to gpkg (only buildings) -- RUN ONLY ONCE 
# osm_no_pbf <- "data/big_data/OSM/norway-latest.osm.pbf"
# osm_no_building_gpkg <- osmextract::oe_vectortranslate(osm_pbf,
#                                                 layer = "multipolygons")

# # Transform OSM data to gpkg (only buildings) -- RUN ONLY ONCE
# osm_se_pbf <- "data/big_data/OSM/sweden-latest.osm.pbf"
# osm_se_building_gpkg <- osmextract::oe_vectortranslate(osm_se_pbf,
#                                                 layer = "multipolygons")

#' Function for getting buildings counts from OSM
#' @description Read only buildings from OSM.gpkg file (by tiles) and calculate building density
#' @param .tile_path Tiles where building counts should be calculated 
#' @param .data_path path to folder with OSM data (.gpkg format) 
get_building_counts <- function(.tile_path,
                                .data_path = "data/big_data/OSM"){

  # read tile
  tile <- terra::rast(.tile_path)
  files <- list.files(.data_path,
                      pattern="*.gpkg",
                      full.names = TRUE)

  # Get bounding box tile
  wkt_tile <- tile |>
    terra::project("EPSG:4326") |> 
    terra::ext() |> 
    terra::vect() |>
    terra::geom(wkt = TRUE)

  # Calculate number buildings
  bd <- lapply(files, function(.x) { sf::read_sf(.x, wkt_filter = wkt_tile) }) |>  
    dplyr::bind_rows() |> 
    tidyr::drop_na(building) |> 
    dplyr::select(building) |>
    sf::st_transform(crs = 3035) |> 
    sf::st_centroid() |> 
    dplyr::mutate(building = 1) |> 
    terra::vect() |> 
    terra::rasterize(tile, fun = sum)

  return(bd)

}

# Building density ----
build_counts <- region_tiles %>%
  # Density by tiles
  map(get_building_counts) |> 
  # Merge tiles
  sprc() |> 
  mosaic(fun = "mean")
terra::writeRaster(build_counts, 
                   "data/processed_data/build_counts.tif",
                   filetype = "GTiff", 
                   overwrite = TRUE)

# Focal (Smooth values) ----
# Number of buildings in a circle
fwm <- focalMat(build_counts, d = radio_buffer, type = "circle", fillNA = TRUE)
fwm[fwm > 0] <- 1
build_focal <- focal(build_counts, w = fwm, fun = sum , na.rm = TRUE)
terra::writeRaster(build_focal, 
                   "data/processed_data/build_focal.tif",
                   filetype = "GTiff", 
                   overwrite = TRUE)

```

```{r}
#| label: fig-building-density
#| fig-cap: Number of buildings per grid cells of 250 m x 250 m

# Read data
build_counts <- terra::rast("data/processed_data/build_counts.tif")
build_focal <- terra::rast("data/processed_data/build_focal.tif")

# Plot ----
p1 <- ggplot() +
  tidyterra::geom_spatvector(data = region_v, fill = "white") +
  tidyterra::geom_spatraster(data = build_counts) +
  scale_fill_whitebox_c(name = "N. Buildings",
                        palette = "viridi",
                        trans = "log10") +
  labs(title = "Original data from OSM") +
  theme_void()

p2 <- ggplot() +
  tidyterra::geom_spatvector(data = region_v, fill = "white") +
  tidyterra::geom_spatraster(data = build_focal) +
  scale_fill_whitebox_c(name = "N. Buildings",
                        palette = "viridi",
                        trans = "log10") +
   labs(title = "Focal values (circular window)") +
  theme_void()

p1 + p2

```

### Fuel/vegetation maps

#### EFFIS Fuel Map

The EFFIS Fuel Map (@effis_european_2017) is a pan-European map that classifies the vegetation in 10 categories according to the fire behaviour models developed by Anderson, 1982 (@anderson_aids_1982) (i.e., it excludes the last 3 categorise of the model). Resolution 250 x 250 m.

We remove small polygons (lower than 5 km^2^), and calculate the distance from each grid cell to the nearest large wild vegetation. Then, we calculate the grids that are within 2.4 km (`value = 1`) or not (`value = 0`).

We also calculate the percentage of vegetation in the buffer area (`*_wv_per`). In the study region, there are only 8 EFFIS categories (@fig-effis-fuel-map-norway).

```{r}
#| label: fig-effis-fuel-map-norway
#| fig-cap: Wildland vegetation in Norway by type of vegetation (Source EFFIS Fuel Map)
#| cache: true

# # Dowload and unzip data (RUN once)
# effis_fuel_url <- "https://effis-gwis-cms.s3-eu-west-1.amazonaws.com/effis/applications/data-and-services/FuelMap_LAEA.zip"
# dir.create("data/big_data/EFFIS")
# download.file(effis_fuel_url, "data/big_data/EFFIS/FuelMap.zip")
# unzip(zipfile = "data/big_data/EFFIS/FuelMap.zip",
#       exdir = "data/big_data/EFFIS")

# Forest layer norway
effis_fuel_path <- "data/big_data/EFFIS/FuelMap2000_NFFL_LAEA.tif"
effis_fuel <- rast(effis_fuel_path) |> 
  crop(region_v) |> 
  mask(region_v)

# Wildland vegetation
effis_wild_veg <- effis_fuel
effis_wild_veg[effis_wild_veg >= 1] <- 1
names(effis_wild_veg) <- "wild_veg"
terra::writeRaster(effis_wild_veg, 
                   "data/processed_data/effis_wild_veg.tif",
                   filetype = "GTiff", 
                   overwrite = TRUE)

# Large forest areas (> 5km2)
effis_wild_veg_large <- stars::read_stars("data/processed_data/effis_wild_veg.tif") |> 
  get_large_polygons() 
terra::writeRaster(effis_wild_veg_large, 
                   "data/processed_data/effis_wild_veg_large.tif",
                   filetype = "GTiff", 
                   overwrite = TRUE)

# Distances to large vegetated areas ----
effis_wild_veg_large_dist <- terra::distance(effis_wild_veg_large)
terra::writeRaster(effis_wild_veg_large_dist, 
                   "data/processed_data/effis_wild_veg_large_dist.tif",
                   filetype = "GTiff", 
                   overwrite = TRUE)

# Percentage-vegetation 
effis_wv_perc <- get_perc_veg_buffer(
  .veg = effis_wild_veg,
  .radio_buffer = radio_buffer)
terra::writeRaster(effis_wv_perc, 
                   "data/processed_data/effis_wv_perc.tif",
                   filetype = "GTiff", 
                   overwrite = TRUE)

# Plot ----
effis_wild_veg_list_no <- tribble(~code, ~cover_class, ~RGB,
                               1, "Short grass",	"16-240-96",
                               3, "Tall grass",	"16-128-16",
                               5, "Brush",	"255-255-0",
                               6, "Dormant brush", "255-204-0",
                               7, "Southern rough",	"255-102-0",
                               8, "Closed timber litter", "192-128-64",
                               9, "Hardwood litter", "160-64-0",
                               10, "Timber", "112-48-48") |>
  separate(RGB, c("red", "green", "blues"), sep = "-") |>  
  mutate(rgb_code = rgb(red, green, blues, maxColorValue = 255))

# Plot
effis_fuel_cls <- effis_fuel
cls <- data.frame(id = effis_wild_veg_list_no$code, 
                  value = effis_wild_veg_list_no$cover_class)
levels(effis_fuel_cls) <- cls

plot(region_r, col = "#E7EEE7", axes = FALSE, legend = FALSE)
plot(effis_fuel_cls, 
     col = effis_wild_veg_list_no$rgb_code,
     axes = FALSE, add = T)

```

```{r}
#| label: fig-effis-wildland-vegetation-norway
#| fig-cap: a) Wildland vegetation in Norway, and b) grid cells within a distance to large vegetation areas (>5km2) of 2.4 km (Source EFFIS Fuel Map)

# Read Effis data
effis_wild_veg <- terra::rast("data/processed_data/effis_wild_veg.tif")

# Get cells within 2400 m (YES --> 1; NO --> 0)
dist_path <- "data/processed_data/effis_wild_veg_large_dist.tif"
effis_wild_veg_large_dist_2400 <- terra::rast(dist_path) |> 
  get_dist_max(2400)

# Plot
par(mfrow = c(1,2), adj = 0)

# Wildand vegetation (Yes = 1, No = NA)
plot(region_r, col = "#E7EEE7", axes = F, legend = F)
plot(effis_wild_veg, col = "#009E73", axes = F, legend = F, add = T)
title("A)", adj = 0)
# Within a distance of 2.4 km (Yes = 1, No = 0)
plot(region_r, col = "#E7EEE7", axes = F, legend = F)
plot(effis_wild_veg_large_dist_2400, col = c(NA, "#0072B2"), axes = F, legend = F, add = T)
title("B)", adj = 0)

```

#### Corine Land Cover

We have used the Corine Land Cover 2018 (Version 2020_20u1), downloaded from [Copernicus](https://land.copernicus.eu/pan-european/corine-land-cover). The resolution of the raster is 100 m x 100 m. We classify as potential wildland vegetation the following classes @fig-corine-wild-vegetation:

-   311: [Broad-leaved forest](https://land.copernicus.eu/user-corner/technical-library/corine-land-cover-nomenclature-guidelines/html/index-clc-311.html)
-   312: [Coniferous forest](https://land.copernicus.eu/user-corner/technical-library/corine-land-cover-nomenclature-guidelines/html/index-clc-312.html)
-   313: [Mixed forest](https://land.copernicus.eu/user-corner/technical-library/corine-land-cover-nomenclature-guidelines/html/index-clc-313.html)
-   321: [Natural grassland](https://land.copernicus.eu/user-corner/technical-library/corine-land-cover-nomenclature-guidelines/html/index-clc-321.html)
-   322: [Moors and heathland](https://land.copernicus.eu/user-corner/technical-library/corine-land-cover-nomenclature-guidelines/html/index-clc-322.html)
-   323: [Sclerophyllous vegetation](https://land.copernicus.eu/user-corner/technical-library/corine-land-cover-nomenclature-guidelines/html/index-clc-323.html)
-   324: [Transitional woodland-shrub](https://land.copernicus.eu/user-corner/technical-library/corine-land-cover-nomenclature-guidelines/html/index-clc-324.html)

```{r}
#| label: fig-corine-fuel-map-norway
#| fig-cap: Wildland vegetation in Norway by type of vegetation (Source CORINE)
#| cache: true

# Load data ----
clc2018_path <- "data/big_data/corine/DATA/U2018_CLC2018_V2020_20u1.tif"

clc2018_r <- rast(clc2018_path) |> 
  crop(region_v) |> 
  mask(region_v)

# Get Wildland vegetation ----
clc2018_wild_veg_list <- tribble(~CODE_18, ~LABEL3, ~RGB,
                                 311,   "Broad-leaved forest",  "128-255-000",
                                 312,   "Coniferous forest",    "000-166-000",
                                 313,   "Mixed forest", "077-255-000",
                                 321,   "Natural grasslands",   "204-242-077",
                                 322,   "Moors and heathland",  "166-255-128",
                                 323,   "Sclerophyllous vegetation", "166-230-077",
                                 324,   "Transitional woodland-shrub",  "166-242-000") |>
  separate(RGB, c("red", "green", "blues"), sep = "-") |>  
  mutate(rgb_code = rgb(red, green, blues, maxColorValue = 255))

# Subset target vegetation
corine_fuel_veg <- clc2018_r |> 
  terra::subst(clc2018_wild_veg_list$LABEL3, clc2018_wild_veg_list$LABEL3, others = NA) |> 
  droplevels()

# Wildland vegetation: yes --> 1 
corine_wild_veg <- as.int(corine_fuel_veg)
corine_wild_veg[corine_wild_veg > 0] <- 1
# Remove colours associated with the rasters
coltab(corine_wild_veg) <- NULL
names(corine_wild_veg) <- "wild_veg"
terra::writeRaster(corine_wild_veg, 
                   "data/processed_data/corine_wild_veg.tif",
                   filetype = "GTiff", 
                   overwrite = TRUE)

# Large forest areas (> 5km2)
corine_wild_veg_large <- stars::read_stars("data/processed_data/corine_wild_veg.tif") |> 
  get_large_polygons() 
terra::writeRaster(corine_wild_veg_large, 
                   "data/processed_data/corine_wild_veg_large.tif",
                   filetype = "GTiff", 
                   overwrite = TRUE)

# Distances to large vegetated areas ----
set.values(corine_wild_veg_large)
corine_wild_veg_large_dist <- terra::distance(corine_wild_veg_large)
terra::writeRaster(corine_wild_veg_large_dist, 
                   "data/processed_data/corine_wild_veg_large_dist.tif",
                   filetype = "GTiff", 
                   overwrite = TRUE)

# Percentage-vegetation ----
corine_wv_perc <- get_perc_veg_buffer(
  .veg = corine_wild_veg,
  .radio_buffer = radio_buffer)
terra::writeRaster(corine_wv_perc, 
                   "data/processed_data/corine_wv_perc.tif",
                   filetype = "GTiff", 
                   overwrite = TRUE)

# Plot ----
plot(region_r, col = "#E7EEE7", axes = FALSE, legend = FALSE)
plot(corine_fuel_veg, col = clc2018_wild_veg_list$rgb_code, 
     axes = FALSE, add = T)

```

```{r}
#| label: fig-wild-vegetation-corine
#| fig-cap: a) Wildland vegetation in Norway, and b) grid cells within a distance to large vegetation areas (>5km2) of 2.4 km (Source CORINE)

# Read CORINE data
corine_wild_veg <- terra::rast("data/processed_data/corine_wild_veg.tif")

# Get cells within 2400 m (YES --> 1; NO --> 0)
dist_path <- "data/processed_data/corine_wild_veg_large_dist.tif"
corine_wild_veg_large_dist_2400 <- terra::rast(dist_path) |> 
  get_dist_max(2400)

# Plot
par(mfrow = c(1,2), adj = 0)

# Wildland vegetation (Yes = 1, No = NA)
plot(region_r, col = "#E7EEE7", axes = F, legend = F)
plot(corine_wild_veg, col = "#009E73", axes = F, legend = F, add = T)
title("A)", adj = 0)
# Within a distance of 2.4 km (Yes = 1, No = 0)
plot(region_r, col = "#E7EEE7", axes = F, legend = F)
plot(corine_wild_veg_large_dist_2400, col = c(NA, "#0072B2"), axes = F, legend = F, add = T)
title("B)", adj = 0)

```

#### National datasets

##### Norwegian Forest resource map - SR16

We downloaded the Norwegian Forest Resource map (Skogressurskart: SR16 - raster format: resolution 16 m) from [NIBIO](https://www.nibio.no/tema/skog/kart-over-skogressurser/skogressurskart-sr16). We downloaded the raster data by county (*Fylke*) in EUREF89 UTM 33, 2d (25833). There are multiple raster files but we only used the dominant tree species (*SRTRESLAG*), which is classified in three categories:

\- *gran* - spruce: 1

\- *furu* - pine: 2

\- *lauv* - deciduous: 3

We consider all of them as wildland vegetation (we reclassified the raster as 1 = wildland vegetation, and 0 = No wildland vegetation).

We aggregate the data by a factor of 6.25, so we past from an original resolution of 16x16 meters to 100x100 meters.

```{r}
#| label: SR16-fuel-map-norway
#| cache: true

# Vegetation layer norway ----
sr16r_path <- list.files("data/big_data/sr16/srtreslag",
                         pattern = "*.tif",
                         full.names = TRUE)
sr16_r <- map(sr16r_path, rast)

# Wildland vegetation
get_wild_veg <- function(.x){
    r <- .x
    r[r >= 1] <- 1 
    return(r)
  }
sr16_wild_veg <-  map(sr16_r, get_wild_veg)

# Aggregate 
change_resolution <- function(.x, .fact = 6.25) {
  aggregate(.x, fact = .fact, fun = "modal") 
}

sr16_wild_veg <- map(sr16_wild_veg, change_resolution)

# Merge rasters
sr16_wild_veg <- sr16_wild_veg |> 
  sprc() |> 
  mosaic(fun = "max") |> 
  project("EPSG:3035", method = "near")

# Export as .tif
set.values(sr16_wild_veg)
names(sr16_wild_veg) <- "wild_veg"
terra::writeRaster(sr16_wild_veg,
                   "data/processed_data/sr16_wild_veg.tif", 
                   filetype = "GTiff", 
                   overwrite = TRUE)

# Large forest areas (> 5km2)
sr16_wild_veg_large <- stars::read_stars("data/processed_data/sr16_wild_veg.tif") |>   get_large_polygons() 
set.values(sr16_wild_veg_large)
terra::writeRaster(sr16_wild_veg_large, 
                   "data/processed_data/sr16_wild_veg_large.tif",
                   filetype = "GTiff", 
                   overwrite = TRUE)

# Distances to large vegetated areas ----
sr16_wild_veg_large_dist <- terra::distance(sr16_wild_veg_large)
set.values(sr16_wild_veg_large_dist)
terra::writeRaster(sr16_wild_veg_large_dist, 
                   "data/processed_data/sr16_wild_veg_large_dist.tif",
                   filetype = "GTiff", 
                   overwrite = TRUE)

# Percentage-vegetation ----
sr16_wv_perc <- get_perc_veg_buffer(
  .veg = sr16_wild_veg,
  .radio_buffer = radio_buffer)
terra::writeRaster(sr16_wv_perc, 
                   "data/processed_data/sr16_wv_perc.tif",
                   filetype = "GTiff", 
                   overwrite = TRUE)

```

```{r}
#| label: fig-SR16-fuel-map-norway
#| fig-cap: Wildland vegetation in Norway by type of vegetation (Source SR16)

# Plot original data ----
categorical_levels <- function(.x){
  r <- .x
  levels(r) <- data.frame(id = 1:3, type =  c("spruce", "pine", "deciduous"))
  return(r)
}

sr16_vrt <- terra::vrt(sr16r_path) 
levels(sr16_vrt) <- data.frame(id = 1:3, type =  c("spruce", "pine", "deciduous"))
no_v_p <- project(no_v, "EPSG:25833")
plot(no_v_p, col = "#E7EEE7", axes = FALSE, legend = FALSE)
plot(sr16_vrt, col = c("lightgreen", "darkgreen", "yellow"), axes = FALSE, add = T)

```

##### Swedish National Land Cover Database (NMD)

We download the NMD Base map from the Swedish Environmental Protection Agency ([Nationella marktäckedata 2018; basskikt](https://geodata.naturvardsverket.se/nedladdning/marktacke/NMD2018/NMD2018_basskikt_ogeneraliserad_Sverige_v1_1.zip)). It classified Sweden in 6 main levels, which later are subdivided in level 2 and level 3 (@fig-NMD-veg-map-sewden):

+----------------------+----------------------------------------------------+--------------------------------------------------------+
| Level 1              | Level 2                                            | Level 3                                                |
+======================+====================================================+========================================================+
| 1.  Skog             | 1.1 Skog utanför våtmark                           | 1.1.1 Tallskog utanför våtmark                         |
|                      |                                                    |                                                        |
|                      |                                                    | 1.1.2 Granskog utanför våtmark                         |
|                      |                                                    |                                                        |
|                      |                                                    | 1.1.3 Barrblandskog utanför våtmark                    |
|                      |                                                    |                                                        |
|                      |                                                    | 1.1.4 Lövblandad barrskog utanför våtmark              |
|                      |                                                    |                                                        |
|                      |                                                    | 1.1.5 Triviallövskog utanför våtmark                   |
|                      |                                                    |                                                        |
|                      |                                                    | 1.1.6 Ädellövskog utanför våtmark                      |
|                      |                                                    |                                                        |
|                      |                                                    | 1.1.7 Triviallövskog med ädellövinslag utanför våtmark |
|                      |                                                    |                                                        |
|                      |                                                    | 1.1.8 Temporärt ej skog utanför våtmark                |
+----------------------+----------------------------------------------------+--------------------------------------------------------+
|                      | 1.2 Skog på våtmark                                | 1.2.1 Tallskog på våtmark                              |
|                      |                                                    |                                                        |
|                      |                                                    | 1.2.2 Granskog på våtmark                              |
|                      |                                                    |                                                        |
|                      |                                                    | 1.2.3 Barrblandskog på våtmark                         |
|                      |                                                    |                                                        |
|                      |                                                    | 1.2.4 Lövblandad barrskog på våtmark                   |
|                      |                                                    |                                                        |
|                      |                                                    | 1.2.5 Triviallövskog på våtmark                        |
|                      |                                                    |                                                        |
|                      |                                                    | 1.2.6 Ädellövskog på våtmark                           |
|                      |                                                    |                                                        |
|                      |                                                    | 1.2.7 Triviallövskog med ädellövinslag på våtmark      |
|                      |                                                    |                                                        |
|                      |                                                    | 1.2.8 Temporärt ej skog på våtmark                     |
+----------------------+----------------------------------------------------+--------------------------------------------------------+
| 2.  Våtmark          |                                                    |                                                        |
+----------------------+----------------------------------------------------+--------------------------------------------------------+
| 3.  Åkermark         |                                                    |                                                        |
+----------------------+----------------------------------------------------+--------------------------------------------------------+
| 4.  Övrig öppen mark | 4.1 Övrig öppen mark utan vegetation               |                                                        |
+----------------------+----------------------------------------------------+--------------------------------------------------------+
|                      | 4.2 Övrig öppen mark med vegetation                |                                                        |
+----------------------+----------------------------------------------------+--------------------------------------------------------+
| 5.  Exploaterad mark | 5.1 Exploaterad mark, byggnad                      |                                                        |
+----------------------+----------------------------------------------------+--------------------------------------------------------+
|                      | 5.2 Exploaterad mark, ej byggnad eller väg/järnväg |                                                        |
+----------------------+----------------------------------------------------+--------------------------------------------------------+
|                      | 5.3 Exploaterad mark, väg/järnväg                  |                                                        |
+----------------------+----------------------------------------------------+--------------------------------------------------------+
| 6.  Vatten           | 6.1 Sjö och vattendrag                             |                                                        |
+----------------------+----------------------------------------------------+--------------------------------------------------------+
|                      | 6.2 Hav                                            |                                                        |
+----------------------+----------------------------------------------------+--------------------------------------------------------+

: NMD categories (table Bilaga 1: Nomenklatur)

```{r}
#| label: fig-NMD-veg-map-sewden
#| fig-cap: Vegetation map of Sweden (National Land Cover Database 2018)

nmd_path <- "data/big_data/NMD2018/nmd2018bas_ogeneraliserad_v1_1.tif" 
nmd_r <- rast(nmd_path)

# Vegetation code
levles_lst <- levels(nmd_r)
levles_lst[[1]]$Klass <- iconv(levles_lst[[1]]$Klass, "latin1")
levels(nmd_r) <- levles_lst

nmd_veg_list <- levles_lst[[1]] |>
  as_tibble() |> 
  mutate(Klass = ifelse(Klass == "", NA, Klass)) |> 
  drop_na()

plot(nmd_r)

```

Then, we classified as Non-burnable the following categories:

```{r}
#| label: tbl-nmd-burnable
#| tbl-cap: Non-burnable vegetation (categories)

nmd_no_fuel_list <- nmd_veg_list |> 
  filter(value %in% c(3, 41, 42, 51, 52, 53, 61, 62, 131, 169, 170, 179, 181, 189, 190))

nmd_no_fuel_list |> 
  gt()

```

The original resolution of the map is 10 meters. We aggregate it (by a factor of 10) and transform the coordinate system from `SWEREF99 TM (EPSG:3006)` to `ETRS89-extended / LAEA Europe (EPSG:3035)`.

```{r}
#| label: NMD-wild-veg-map-sewden
#| cache: true

# Reduce resolution 
nmd_r_a <- aggregate(nmd_r, fact = 10, fun = "modal") |> 
  project("EPSG:3035", method = "near")

# Remove no fuel vegetations 
nmd_wild_veg <- nmd_r_a |> 
  mutate(Klass = ifelse(Klass %in% nmd_no_fuel_list$Klass, NA, Klass))

# Wild vegetation; YES = 1, NO = 0
nmd_wild_veg[nmd_wild_veg == 1] <- NA  # Outside Sweden 
nmd_wild_veg[nmd_wild_veg  > 1] <- 1

# Export as .tif
names(nmd_wild_veg) <- "wild_veg"
coltab(nmd_wild_veg) <- NULL
set.values(nmd_wild_veg)
terra::writeRaster(nmd_wild_veg,
                   "data/processed_data/nmd_wild_veg.tif", 
                   filetype = "GTiff", 
                   overwrite = TRUE)

# Large forest areas (> 5km2)
nmd_wild_veg_large <- stars::read_stars("data/processed_data/nmd_wild_veg.tif") |> 
  get_large_polygons() 
set.values(nmd_wild_veg_large)
terra::writeRaster(nmd_wild_veg_large, 
                   "data/processed_data/nmd_wild_veg_large.tif",
                   filetype = "GTiff", 
                   overwrite = TRUE)

# Distances to large vegetated areas ----
nmd_wild_veg_large_dist <- terra::distance(nmd_wild_veg_large)
set.values(nmd_wild_veg_large_dist)
terra::writeRaster(nmd_wild_veg_large_dist, 
                   "data/processed_data/nmd_wild_veg_large_dist.tif",
                   filetype = "GTiff", 
                   overwrite = TRUE)

# Percentage-vegetation ----
nmd_wv_perc <- get_perc_veg_buffer(
  .veg = nmd_wild_veg,
  .radio_buffer = radio_buffer)
set.values(nmd_wv_perc)
terra::writeRaster(nmd_wv_perc, 
                   "data/processed_data/nmd_wv_perc.tif",
                   filetype = "GTiff", 
                   overwrite = TRUE)

```

Finally, we combine both maps in one (@fig-wild-vegetation-national)

```{r}
#| label: fig-wild-vegetation-national
#| fig-cap: a) Wildland vegetation in the region, and b) grid cells within a distance to large vegetation areas (>5km2) of 2.4 km (Source SR16 and NMD)

# Read SR16 data
sr16_wild_veg <- terra::rast("data/processed_data/sr16_wild_veg.tif")
dist_path <- "data/processed_data/sr16_wild_veg_large_dist.tif"
sr16_wild_veg_large_dist_2400 <- terra::rast(dist_path) |> 
  get_dist_max(2400)

# Read NRM data
nmd_wild_veg <- terra::rast("data/processed_data/nmd_wild_veg.tif")
dist_path <- "data/processed_data/nmd_wild_veg_large_dist.tif"
nmd_wild_veg_large_dist_2400 <- terra::rast(dist_path) |> 
  get_dist_max(2400)

# Plot
par(mfrow = c(1,2), adj = 0)

# Wildland vegetation (Yes = 1, No = NA)
plot(region_r, col = "#E7EEE7", axes = F, legend = F)
plot(sr16_wild_veg, col = "#009E73", axes = F, legend = F, add = T)
plot(nmd_wild_veg, col = "#009E73", axes = F, legend = F, add = T)
title("A)", adj = 0)
# Within a distance of 2.4 km (Yes = 1, No = 0)
plot(region_r, col = "#E7EEE7", axes = F, legend = F)
plot(sr16_wild_veg_large_dist_2400, 
     col = c(NA, "#0072B2"),
     axes = F, 
     legend = F,
     add = T)
plot(nmd_wild_veg_large_dist_2400, 
     col = c(NA, "#0072B2"), 
     axes = F, 
     legend = F, 
     add = T)
title("B)", adj = 0)

```

#### FirEUrisk_Europe_fuel_map

Original data from [@Aragoneses2022]. Al fuels are considered for WUI. The resolution of the original map is 1km and we have re-sampled it to a grid cell of 250x250 grids (same than `region_r`) for the WUI (based in the type that is most presented in the area: `mode`).

```{r}
#| label: fig-firEUrisk-fuel-map-norway
#| fig-cap: Wildland vegetation in Norway by type of vegetation (Source EFFIS Fuel Map)
#| cache: true

# Codes
firEUrisk_fuel_type <- tribble(~code, ~description,
                                1111,"Open broadleaf evergreen forest",
                                  23, "High shrubland",
                                1112, "Closed broadleaf evergreen forest",
                                  31, "Low grassland",
                                1121, "Open broadleaf deciduous forest",
                                  32, "Medium grassland",
                                1122, "Closed broadleaf deciduous forest",
                                  33, "High grassland",
                                1211, "Open needleleaf evergreen forest",
                                  41, "Herbaceous cropland",
                                1212, "Closed needleleaf evergreen forest",
                                  42, "Woody cropland",
                                1221, "Open needleleaf deciduous forest",
                                  51, "Wet and peat/semi-peat land – tree",
                                1222, "Closed needleleaf deciduous forest", 
                                  52, "Wet and peat/semi-peat land – shrubland",
                                1301, "Open mixed forest",
                                  53, "Wet and peat/semi-peat land – grassland",
                                1302, "Closed mixed forest",
                                  61, "Urban continuous fabric",
                                  21, "Low shrubland",
                                  62, "Urban discontinuous fabric",
                                  22, "Medium shrubland",
                                   7, "Nonfuel")

firEUrisk_no_fuel_type <- tribble(~code, ~description,
                                     41, "Herbaceous cropland",
                                     42, "Woody cropland",
                                     51, "Wet and peat/semi-peat land – tree",
                                     52, "Wet and peat/semi-peat land – shrubland",
                                     53, "Wet and peat/semi-peat land – grassland",
                                     61, "Urban continuous fabric",
                                     62, "Urban discontinuous fabric",
                                      7, "Nonfuel")

# Forest layer norway
firEUrisk_fuel_path <- "data/big_data/firEUrisk/FirEUrisk_Europe_fuel_map.tif"
firEUrisk_fuel <- rast(firEUrisk_fuel_path) |> 
  crop(region_v) |> 
  mask(region_v)

# Resample to grid cells of 250m x 250m (same than region_r)
firEUrisk_fuel_250m <- resample(firEUrisk_fuel, region_r, method = "mode") |> 
 # Remove no fuel vegetations 
  mutate(FirEUrisk_Europe_fuel_map = ifelse(FirEUrisk_Europe_fuel_map %in% firEUrisk_no_fuel_type$code,
                    NA,
                    FirEUrisk_Europe_fuel_map))

# Wildland vegetation
firEUrisk_wild_veg <- firEUrisk_fuel_250m 
firEUrisk_wild_veg[firEUrisk_wild_veg >= 1] <- 1
names(firEUrisk_wild_veg) <- "wild_veg"
terra::writeRaster(firEUrisk_wild_veg, 
                   "data/processed_data/firEUrisk_wild_veg.tif",
                   filetype = "GTiff", 
                   overwrite = TRUE)

# Large forest areas (> 5km2)
firEUrisk_wild_veg_large <- stars::read_stars("data/processed_data/firEUrisk_wild_veg.tif") |> 
  get_large_polygons() 
terra::writeRaster(firEUrisk_wild_veg_large, 
                   "data/processed_data/firEUrisk_wild_veg_large.tif",
                   filetype = "GTiff", 
                   overwrite = TRUE)

# Distances to large vegetated areas ----
set.values(firEUrisk_wild_veg_large)
firEUrisk_wild_veg_large_dist <- terra::distance(firEUrisk_wild_veg_large)
terra::writeRaster(firEUrisk_wild_veg_large_dist, 
                   "data/processed_data/firEUrisk_wild_veg_large_dist.tif",
                   filetype = "GTiff", 
                   overwrite = TRUE)

# Percentage-vegetation ----
firEUrisk_wv_perc <- get_perc_veg_buffer(
  .veg = firEUrisk_wild_veg,
  .radio_buffer = radio_buffer)
terra::writeRaster(firEUrisk_wv_perc, 
                   "data/processed_data/firEUrisk_wv_perc.tif",
                   filetype = "GTiff", 
                   overwrite = TRUE)
# Plot ----
levels(firEUrisk_fuel_250m) <- data.frame(id = firEUrisk_fuel_type$code,
                                          type =  firEUrisk_fuel_type$description)

plot(region_r, col = "#E7EEE7", axes = FALSE, legend = FALSE)
plot(firEUrisk_fuel_250m, axes = FALSE, add = T)

```

```{r}
#| label: fig-firEUrisk-wildland-vegetation-norway
#| fig-cap: a) Wildland vegetation in Norway, and b) grid cells within a distance to large vegetation areas (>5km2) of 2.4 km (Source EFFIS Fuel Map)

# Read Effis data
firEUrisk_wild_veg <- terra::rast("data/processed_data/firEUrisk_wild_veg.tif")
dist_path <- "data/processed_data/firEUrisk_wild_veg_large_dist.tif"
firEUrisk_wild_veg_large_dist_2400 <- terra::rast(dist_path) |> 
  get_dist_max(2400)

# Plot
par(mfrow = c(1,2), adj = 0)

# Wildand vegetation (Yes = 1, No = NA)
plot(region_r, col = "#E7EEE7", axes = F, legend = F)
plot(firEUrisk_wild_veg, col = "#009E73", axes = F, legend = F, add = T)
title("A)", adj = 0)
# Within a distance of 2.4 km (Yes = 1, No = 0)
plot(region_r, col = "#E7EEE7", axes = F, legend = F)
plot(firEUrisk_wild_veg_large_dist_2400, col = c(NA, "#0072B2"), axes = F, legend = F, add = T)
title("B)", adj = 0)

```

## WUI classification

We have used the definition described by [@stewart_defining_2007; @bar-massada_comparative_2021; @bar-massada_using_2013], which is shown in [@fig-WUI-raster; @bar-massada_comparative_2021]:

```{r}
#| label: fig-WUI-raster 
#| fig-cap: "Zonal-based WUI mapping approach"

knitr::include_graphics("figures/wui_map_approach_bar_massada_2021.png")

```

We carry out a point-based approach with grid cells of `r raster_resolution` m x `r raster_resolution` m, and a buffer radio of `r radio_buffer` m. Our workflow is therefore:

```{mermaid}
flowchart TD
  A[Grid 250m x 250m] --> B{"Building density > 6.17 #/km2"}
  B --  Yes  --> C{Vegetation > 50%}
  B --  No   --> D[No WUI]
  C --  Yes  --> E[Intermix] 
  C --  No   --> F{within 2.4 km of large forests (>5 km2)}
  F --  Yes  --> G[Interface]
  F --  No   --> H[No WUI]
  style E fill:#CC79A7
  style G fill:#F0E442
  style D fill:#D3D3D3
  style H fill:#D3D3D3
```

Therefore, we do the following steps:

1.  Create a raster with the potential WUI cells (`wui_potential`)

-   Building density (in the buffer area) ≤ 6.16 #/km\^2 --\> NA

-   Building density (in the buffer area) \> 6.16 #/km\^2 --\> 1

2.  Create a raster with the grids where the percentage of fuel vegetation in the buffer area is higher than 50% (`wv_perc`)

-   Percentage \< 50% --\> 0
-   Percentage ≥ 50% --\> 1

3.  Create a raster indicating if the grid cell is within 2.4 km of a large (>5 km2) fuel vegetation area (`wild_veg_large_dist_*`)

-   Yes: 1
-   No: 0

4.  Sum both rasters (`wv_perc + wui_potential`), and generate a new raster (`wui`) with three options :

-   wui = NA --\> No WUI

-   wui = 2 --\> WUI intermix

-   wui = 1 --\> Potential interface (it is reassigned to 0 in the raster)

5.  Sum this raster (`wui`) and `wild_veg_large_dist_`, we get therefore that:

-   wui = NA --\> No WUI
-   wui = 0 --\> No WUI
-   wui = 1 --\> Interface
-   wui = 2 --\> Intermix
-   wui = 3 --\> Intermix

```{r}
#| label: fig-potential-wui
#| fig-cap: Potential WUI grids based on building density

# Minimum number of buildings (N) in the buffer area ----
# WUI if building density (d) > 6.17 [#/km^2]
# N = d [#/km^2] * A [km^2]
# A = pi * radio^2
d_min <- 6.17
N_min <- round(d_min * area_buffer, 3)

wui_potential <- build_focal
wui_potential[wui_potential <= N_min] <- NA
wui_potential[wui_potential >  N_min] <- 1

plot(region_r, col = "#E7EEE7", axes = F, legend = F)
plot(wui_potential, axes = F, legend = F, add = T)

```

```{r}
#| label: read-wv-data

effis_wv_perc     <- terra::rast("data/processed_data/effis_wv_perc.tif")
corine_wv_perc    <- terra::rast("data/processed_data/corine_wv_perc.tif")
sr16_wv_perc      <- terra::rast("data/processed_data/sr16_wv_perc.tif")
nmd_wv_perc       <- terra::rast("data/processed_data/nmd_wv_perc.tif")
firEUrisk_wv_perc <- terra::rast("data/processed_data/firEUrisk_wv_perc.tif")

```

```{r}
#| label: calculate-wui-region
#| cache: true

# EFFIS
effis_wui <- get_wui_raster(.veg_perc = effis_wv_perc,
                            .veg_dist = effis_wild_veg_large_dist_2400)

# CORINE
corine_wui <- get_wui_raster(.veg_perc = corine_wv_perc,
                             .veg_dist = corine_wild_veg_large_dist_2400)

# National maps
sr16_wui <- get_wui_raster(.veg_perc = sr16_wv_perc,
                           .veg_dist = sr16_wild_veg_large_dist_2400,
                           .region = no_v)

nrm_wui <- get_wui_raster(.veg_perc = nmd_wv_perc,
                          .veg_dist = nmd_wild_veg_large_dist_2400,
                          .region = se_v)

national_wui <- mosaic(sr16_wui, nrm_wui, fun = "max") |> 
  add_colours_wui()

# firEUrisk
firEUrisk_wui <- get_wui_raster(.veg_perc = firEUrisk_wv_perc,
                                .veg_dist = firEUrisk_wild_veg_large_dist_2400)

# Put all rasters in one 
wui_r <- c(effis_wui, corine_wui, national_wui, firEUrisk_wui)
names(wui_r) <- c("effis", "corine", "national", "firEUrisk")

terra::saveRDS(wui_r, "data/processed_data/wui_r.rds")

```

Map of WUI areas (@fig-wui-maps-region; @tbl-pct-wui-region).

```{r}
#| label: read-wui-raster

wui_r <- terra::readRDS("data/processed_data/wui_r.rds")

```

```{r}
#| label: fig-wui-maps-region
#| fig-cap: "WUI maps of Norway based on different fuel/vegetation maps (resolution of 250 x 250 m)"
#| fig-subcap: 
#|   - "EFFIS"
#|   - "CORINE"
#|   - "FirEUrisk"
#|   - "National databases"
#| layout-ncol: 2
#| fig-width: 4
#| fig-format: png
#| cache: true

f_plot <- function(.x){ plot(.x, 
                             axes = F, 
                             mar = c(0, 0, 0, 0),
                             legend = "topleft")  }

f_plot(wui_r$effis)
f_plot(wui_r$corine)
f_plot(wui_r$firEUrisk)
f_plot(wui_r$national)

```

Percentages at national scale from the different vegetation/fuel types (based on the number of grids cells of `r raster_resolution`x `r raster_resolution` \[m\] for each WUI type):

```{r}
#| label: tbl-pct-wui-region
#| tbl-cap: Percentages of WUI by vegetation and country
#| tbl-subcap: 
#|  - Norway
#|  - Sweden
#| layout-ncol: 2

# summary table 
tbl_summary_vegetation <- function(.x){

  .x |> 
    as_tibble() |> 
    drop_na() |> 
    pivot_longer(everything(), 
                 names_to = "name",
                 values_to = "value") |> 
    group_by(name, value) |> 
    summarise(count = n()) |> 
    ungroup() |> 
    group_by(name) |> 
    mutate(total = sum(count),
           pct = 100*count/total) |> 
    pivot_wider(names_from = "value", 
                values_from = c("count", "pct")) |> 
    mutate(pct_wui = pct_Interface + pct_Intermix) |> 
    select(starts_with("pct")) |> 
    relocate(pct_wui, .after = pct_No_wui) |> 
    gt(row_group_as_column = TRUE) |> 
    fmt_number(decimals = 1) |> 
    cols_label(
      name = "",
      pct_No_wui = md("**No WUI**"),
      pct_Interface = md("**Interface**"),
      pct_Intermix = md("**Intermix**"),
      pct_wui = md("**Total**")
    )  |> 
    tab_spanner(
      label = md("**WUI**"),
      columns = c(pct_Interface, pct_Intermix, pct_wui)
    ) |> 
    tab_options(
      table.font.size = 9,
      row_group.padding = px(1),
      row_group.font.weight = "bold",
      data_row.padding = px(1)
    )

}

# Summary WUI by country 
wui_r_no <- mask(wui_r, no_v)
wui_r_se <- mask(wui_r, se_v)

# Tables
tbl_summary_vegetation(wui_r_no) 
tbl_summary_vegetation(wui_r_se)

```

Zoom to counties (@fig-wui-map-counties; @tbl-summary-wui-county):

```{r}
#| label: fig-wui-map-counties
#| fig-cap: "WUI maps by county based on different fuel/vegetation maps (resolution of 250 x 250 m)"
#| fig-subcap: 
#|  - "Viken"
#|  - "Rogaland"
#|  - "Vestland"
#|  - "Trøndelag"
#|  - "Troms og Finnmark"
#|  - "Innlandet"
#|  - "Oslo"
#|  - "Vestfold og Telemark"
#|  - "Agder"
#|  - "Møre og Romsdal"
#|  - "Nordland"
#| layout-nrow: 11
#| out-width: 100%

# plot_zoom <- function(.x) {
#   
#   zoom_area <- counties |> 
#     filter(NUTS_NAME == {{ .x }})
#   
#   wui_r_zoom <- wui_r |> 
#     crop(zoom_area) |> 
#     mask(zoom_area)
#   
#   # Add colour table 
#   coltab(wui_r_zoom) <- 
#     data.frame(id = 0:2, 
#                col = c("#E7EEE7", "#F0E442", "#CC79A7"))
#   
#   par(mfrow = c(1, 4))
#   
#   plot(wui_r_zoom$effis, 
#        axes = FALSE, 
#        font.main = 1,
#        legend = FALSE,
#        mar = c(0, 0, 0, 0),
#        main =  names(wui_r_zoom$effis))
#   # title( {{ .x }}, line = -4, adj = 0)
#   
#   plot(wui_r_zoom$corine, 
#        axes = FALSE, 
#        font.main = 1,
#        legend = FALSE,
#        mar = c(0, 0, 0, 0),
#        main =  names(wui_r_zoom$corine))
#   
#   plot(wui_r_zoom$firEUrisk,
#        axes = FALSE, 
#        font.main = 1,
#        legend = FALSE,
#        mar = c(0, 0, 0, 0), 
#        main =  names(wui_r_zoom$firEUrisk))
#   
#   plot(wui_r_zoom$sr16,
#        axes = F, 
#        font.main = 1,
#        legend = FALSE,
#        mar = c(0, 0, 0, 0), 
#        main =  names(wui_r_zoom$sr16)) 
#   
# }
# 
# # par(mfrow = c(1, 11))
# as.list(counties$NUTS_NAME) |> 
#   walk(plot_zoom)

```

```{r}
#| label: tbl-summary-wui-county
#| tbl-cap: Percentages of WUI by county
#| tbl-subcap: 
#|  - Norway
#|  - Sweden
#| layout-ncol: 2

# Function for printing the table
tbl_summary_county <- function(.x) {
  .x |> 
    select(NUTS_NAME, name, starts_with("pct")) |> 
    gt(rowname_col = "name",
       groupname_col = "NUTS_NAME",
       row_group_as_column = FALSE) |> 
    fmt_number(decimals = 1) |> 
    cols_label(
      name = "",
      pct_No_wui = md("**No WUI**"),
      pct_Interface = md("**Interface**"),
      pct_Intermix = md("**Intermix**"),
      pct_wui = md("**Total**")
    )  |> 
    tab_spanner(
      label = md("**WUI**"),
      columns = c(pct_Interface, pct_Intermix, pct_wui)
    ) |> 
    tab_options(
      table.font.size = 9,
      row_group.background.color = "#E7EEE7",
      row_group.padding = px(1),
      row_group.font.weight = "bold",
      data_row.padding = px(1)
    ) 
}

# Summary by county
tbl_wui_county <- terra::extract(wui_r, counties) |>  
  pivot_longer(cols = -ID, 
               names_to = "name",
               values_to = "value") |> 
  group_by(ID, name, value) |> 
  summarise(count = n()) |> 
  ungroup() |> 
  group_by(ID, name) |> 
  mutate(total = sum(count), 
         pct = 100*count/total) |> 
  ungroup() |> 
  pivot_wider(names_from = "value",
              values_from = c("count", "pct")) |> 
  mutate(pct_wui = pct_Interface + pct_Intermix)

# Add values to sf counties
counties_wui <- counties |> 
  st_as_sf()  %>% 
  mutate(ID := seq_len(nrow(.)))  |> 
  left_join(tbl_wui_county, by = "ID") |> 
  st_drop_geometry() |> 
  as_tibble()

# Table by country & county 
filter(counties_wui, CNTR_CODE == "NO") |> 
  tbl_summary_county()

filter(counties_wui, CNTR_CODE == "SE") |> 
  tbl_summary_county()

```

Area of wild vegetation in each county: Number vegetation cells (centroid is inside the polygon), sum then and calculate the area based on the raster resolution

```{r}
#| label: tbl-summary-table-areas
#| tbl-cap: Number of buildings and wildland vegetation percentage by county 
#| cache: true

# Number of buildings (use building_count?) ----
county_build_count <- build_counts |> 
  # Number of buildings per county
  extract(y = vect(counties),
          fun = "sum",
          na.rm = TRUE,
          bind = TRUE) |> 
  # sf object
  st_as_sf() %>% 
  # Area polygons
  mutate(area = units::set_units(st_area(.), "km2"))  |> 
  # Building density by county
  mutate(build_den = sum/area)

# Percentage vegetation county ----
# EFFIS
effis_county <- effis_wild_veg |> 
  # Number of cells
  extract(y = vect(counties),
          fun = "sum",
          na.rm = TRUE,
          bind = TRUE) |>  
  st_as_sf() |> 
  # Area (number cells x area cells)
  mutate(
    area_km = wild_veg * ((res(effis_wild_veg))[1]/1000 * (res(effis_wild_veg ))[2]/1000)
    ) |> 
  # Get only values 
  st_drop_geometry() |> 
  select(NUTS_NAME, CNTR_CODE, area_km) |> 
  rename(area_km_effis = area_km)

# CORINE
corine_county <- corine_wild_veg |> 
  # Number of cells
  extract(y = vect(counties),
          fun = "sum",
          na.rm = TRUE,
          bind = TRUE) |>  
  st_as_sf() |> 
  # Area (number cells x area cells)
  mutate(
    area_km = wild_veg * ((res(corine_wild_veg))[1]/1000 * (res(corine_wild_veg ))[2]/1000)
    ) |> 
  # Get only values 
  st_drop_geometry() |> 
  select(NUTS_NAME, CNTR_CODE, area_km) |> 
  rename(area_km_corine = area_km)

# National
sr16_county <- sr16_wild_veg |> 
  # Number of cells
  extract(y = vect(counties),
          fun = "sum",
          na.rm = TRUE,
          bind = TRUE) |>  
  st_as_sf() |> 
  # Area (number cells x area cells)
  mutate(
    area_km = wild_veg * ((res(sr16_wild_veg))[1]/1000 * (res(sr16_wild_veg ))[2]/1000)
    ) |> 
  # Get only values 
  st_drop_geometry() |> 
  select(NUTS_NAME, CNTR_CODE, area_km) |> 
  filter(CNTR_CODE  == "NO")

nmd_county <- nmd_wild_veg |> 
  # Number of cells
  extract(y = vect(counties),
          fun = "sum",
          na.rm = TRUE,
          bind = TRUE) |>  
  st_as_sf() |> 
  # Area (number cells x area cells)
  mutate(
    area_km = wild_veg * ((res(nmd_wild_veg))[1]/1000 * (res(nmd_wild_veg ))[2]/1000)
         ) |> 
  # Get only values 
  st_drop_geometry() |> 
  select(NUTS_NAME, CNTR_CODE, area_km) |> 
  filter(CNTR_CODE  == "SE")

national_county <- rbind(sr16_county, nmd_county) |> 
  rename(area_km_national = area_km)

# FirEUrisk
firEUrisk_county <- firEUrisk_wild_veg |> 
  # Number of cells
  extract(y = vect(counties),
          fun = "sum",
          na.rm = TRUE,
          bind = TRUE) |>  
  st_as_sf() |> 
  # Area (number cells x area cells)
  mutate(
    area_km = wild_veg * ((res(firEUrisk_wild_veg))[1]/1000 * (res(firEUrisk_wild_veg ))[2]/1000)
         ) |> 
  # Get only values 
  st_drop_geometry() |> 
  select(NUTS_NAME, CNTR_CODE, area_km) |> 
  rename(area_km_firEUrisk = area_km)

# Joint tables
county_veg <- effis_county |> 
  left_join(corine_county) |> 
  left_join(national_county) |> 
  left_join(firEUrisk_county)

# Add vegetation to the summary table ----
df_summary <- county_build_count |> 
  st_drop_geometry() |> 
  left_join(county_veg) |> 
  mutate(across(.cols = starts_with("area_km"),
                .fns = ~ 100 * (.x / as.numeric(area)),
                .names = "perc_{.col}"))

# print table
df_summary |> 
  select(CNTR_CODE, NUTS_NAME, area,
         sum, build_den,
         starts_with("perc")) |>
  mutate(area = as.numeric(area),
         build_den = as.numeric(build_den),
         CNTR_CODE = case_when(CNTR_CODE == "NO" ~ "NORWAY",
                               CNTR_CODE == "SE" ~ "SWEDEN")) |> 
  gt(groupname_col = "CNTR_CODE",
     row_group_as_column = FALSE) |> 
  fmt_number(c(build_den, starts_with("perc")), decimals = 1) |> 
  fmt_number(c(sum, area), decimals = 0) |> 
  tab_options(
      table.font.size = 9,
      row_group.background.color = "lightgrey",
      row_group.padding = px(1),
      row_group.font.weight = "bold",
      data_row.padding = px(1)
    ) |> 
  tab_spanner(
      label = md("**Area covered by vegetation [%]**"),
      columns = starts_with("perc")
    ) |> 
  tab_spanner(
      label = md("**Buildings**"),
      columns = c(sum, build_den)
    ) |> 
  cols_label(
      NUTS_NAME = md("**County**"),
      area = md("**Area [km2]**"),
      sum = md("**Count**"),
      build_den = md("**Density [#/km2]**"),
      perc_area_km_effis = md("**Effis**"),
      perc_area_km_corine = md("**Corine**"),
      perc_area_km_national = md("**National**"),
      perc_area_km_firEUrisk = md("**FirEurisk**")
    )

```

### Effect distance to large forest

```{r}
#| label: functions-wui-by-distance-large-forest
# Distance to large forest areas

# Distance variation
max_distances <- c(200, 500, 1000, 2000, 2400)

# WUIs
f_wui <- function(.x, .y){

  veg_perc <- .x

  get_wui_raster(.veg_perc = veg_perc,
                 .veg_dist = .y,
                 .pot_wui = wui_potential,
                 .region = region_v)

}

# Summary table
tbl_summary_distance <- function(.x){

.x |>
  as_tibble() |>
  drop_na() |>
  pivot_longer(everything(),
               names_to = "name",
               values_to = "value") |>
  group_by(name, value) |>
  summarise(count = n()) |>
  ungroup() |>
  group_by(name) |>
  mutate(total = sum(count),
         pct = 100*count/total) |>
  pivot_wider(names_from = "value",
              values_from = c("count", "pct")) |>
  mutate(pct_wui = pct_Interface + pct_Intermix) |>
  select(starts_with("pct")) |>
  relocate(pct_wui, .after = pct_No_wui) |>
  mutate(name = factor(name,
                       levels = c("d200",
                                  "d500",
                                  "d1000",
                                  "d2000",
                                  "d2400"))) |>
  arrange(name) |>
  gt(row_group_as_column = TRUE) |>
  fmt_number(decimals = 1) |>
  cols_label(
    name = "",
    pct_No_wui = md("**No WUI**"),
    pct_Interface = md("**Interface**"),
    pct_Intermix = md("**Intermix**"),
    pct_wui = md("**Total**")
  )  |>
  tab_spanner(
    label = md("**WUI**"),
    columns = c(pct_Interface, pct_Intermix, pct_wui)
  ) |>
  tab_options(
    table.font.size = 9,
    row_group.padding = px(1),
    row_group.font.weight = "bold",
    data_row.padding = px(1)
  )

}

```

#### Effis

```{r}
#| label: fig-wui-effis-vegetations
#| fig-cap: WUI change based on EFFIS fuel map  
#| cache: true

# Raster with distances to large forest (in m)
dist_path <- "data/processed_data/effis_wild_veg_large_dist.tif"
effis_wild_veg_large_dist <- rast(dist_path)

# Withing the max. distance
effis_dist_variation<- map2(effis_wild_veg_large_dist, max_distances, get_dist_max)
names(effis_dist_variation) <- c("d200", "d500", "d1000", "d2000", "d2400")

# Wui vs distance
wui_effis <- map2(effis_wv_perc,
                  effis_dist_variation,
                  f_wui)
wui_effis <- rast(wui_effis)
names(wui_effis) <- c("d200", "d500", "d1000", "d2000", "d2400")
set.values(wui_effis)
terra::saveRDS(wui_effis, "data/processed_data/wui_effis.rsd")

# Plot
panel(wui_effis, axes = FALSE)

```

```{r}
#| label: wui-effis-dist

wui_effis <- terra::readRDS("data/processed_data/wui_effis.rsd")

# Data by country
wui_effis_no <- mask(wui_effis, no_v)
wui_effis_se <- mask(wui_effis, se_v)

```

```{r}
#| label: tbl-wui-effis-dist
#| tbl-cap: Changes in WUI percentages of WUI by country (EFFIS)
#| tbl-subcap: 
#|  - Norway
#|  - Sweden
#| layout-ncol: 2

tbl_summary_distance(wui_effis_no)
tbl_summary_distance(wui_effis_se)

```

#### Corine

```{r}
#| label: fig-wui-corine-vegetations
#| fig-cap: WUI change based on CORINE vegetation map  
#| cache: true

# Raster with distances to large forest (in m)
dist_path <- "data/processed_data/corine_wild_veg_large_dist.tif"
corine_wild_veg_large_dist <- rast(dist_path)

# Withing the max. distance
corine_dist_variation<- map2(corine_wild_veg_large_dist, max_distances, get_dist_max)
names(corine_dist_variation) <- c("d200", "d500", "d1000", "d2000", "d2400")

# Wui vs distance
wui_corine <- map2(corine_wv_perc,
                   corine_dist_variation,
                   f_wui)
wui_corine <- rast(wui_corine)
names(wui_corine) <- c("d200", "d500", "d1000", "d2000", "d2400")
terra::saveRDS(wui_corine, "data/processed_data/wui_corine.rsd")

# Plot
panel(wui_corine, axes = FALSE)

```

```{r}
#| label: wui-corine-dist

wui_corine <- terra::readRDS("data/processed_data/wui_corine.rsd")

# Data by country
wui_corine_no <- mask(wui_corine, no_v)
wui_corine_se <- mask(wui_corine, se_v)

```

```{r}
#| label: tbl-wui-corine-dist
#| tbl-cap: Changes in WUI percentages of WUI by country (CORINE)
#| tbl-subcap: 
#|  - Norway
#|  - Sweden
#| layout-ncol: 2
#| cache: true

tbl_summary_distance(wui_corine_no)
tbl_summary_distance(wui_corine_se)

```

#### National datasets

```{r}
#| label: fig-wui-national-vegetations-maps
#| fig-cap: WUI change based on national vegetation maps  
#| cache: true

# Norway ----
dist_path <- "data/processed_data/sr16_wild_veg_large_dist.tif"
sr16_wild_veg_large_dist <- rast(dist_path)
sr16_dist_variation<- map2(sr16_wild_veg_large_dist, max_distances, get_dist_max)
names(sr16_dist_variation) <- c("d200", "d500", "d1000", "d2000", "d2400")
wui_sr16 <- map2(sr16_wv_perc, sr16_dist_variation, f_wui)
wui_sr16 <- rast(wui_sr16)
names(wui_sr16) <- c("d200", "d500", "d1000", "d2000", "d2400")
terra::saveRDS(wui_sr16, "data/processed_data/wui_sr16.rsd")

# Sweden ----
dist_path <- "data/processed_data/nmd_wild_veg_large_dist.tif"
nmd_wild_veg_large_dist <- rast(dist_path)
nmd_dist_variation<- map2(nmd_wild_veg_large_dist, max_distances, get_dist_max)
names(nmd_dist_variation) <- c("d200", "d500", "d1000", "d2000", "d2400")
wui_nmd <- map2(nmd_wv_perc, nmd_dist_variation, f_wui)
wui_nmd <- rast(wui_nmd)
names(wui_nmd) <- c("d200", "d500", "d1000", "d2000", "d2400")
terra::saveRDS(wui_nmd, "data/processed_data/wui_nmd.rsd")

# Merge datasets ----
wui_national <- mosaic(wui_sr16, wui_nmd, fun = "max") |> 
  add_colours_wui()

# Add names
names(wui_national) <- c("d200", "d500", "d1000", "d2000", "d2400")
set.values(wui_national)
# Export
terra::saveRDS(wui_national, "data/processed_data/wui_national.rsd")

# Plot region  
panel(wui_national, axes = FALSE)

```

```{r}
#| label: wui-national-dist
#| tbl-cap: Changes in WUI percentages of WUI by country (national datasets - SR16 & NMD)

wui_national <- terra::readRDS("data/processed_data/wui_national.rsd")

# Data by country
wui_national_no <- mask(wui_national, no_v)
wui_national_se <- mask(wui_national, se_v)

```

```{r}
#| label: tbl-wui-national-dist
#| tbl-cap: Changes in WUI percentages of WUI by country (national datasets - SR16 & NMD)
#| tbl-subcap: 
#|  - Norway
#|  - Sweden
#| layout-ncol: 2
#| cache: true

tbl_summary_distance(wui_national_no)
tbl_summary_distance(wui_national_se)

```

#### FirEUrisk

```{r}
#| label: fig-wui-firEUrisk-vegetations
#| fig-cap: WUI change based on firEUrisk fuel map  
#| cache: true

# Raster with distances to large forest (in m)
dist_path <- "data/processed_data/firEUrisk_wild_veg_large_dist.tif"
firEUrisk_wild_veg_large_dist <- rast(dist_path)

# Withing the max. distance
firEUrisk_dist_variation<- map2(firEUrisk_wild_veg_large_dist, max_distances, get_dist_max)
names(firEUrisk_dist_variation) <- c("d200", "d500", "d1000", "d2000", "d2400")

# Wui vs distance
wui_firEUrisk <- map2(firEUrisk_wv_perc,
                   firEUrisk_dist_variation,
                   f_wui)
wui_firEUrisk <- rast(wui_firEUrisk)
names(wui_firEUrisk) <- c("d200", "d500", "d1000", "d2000", "d2400")
terra::saveRDS(wui_firEUrisk, "data/processed_data/wui_firEUrisk.rsd")

# Plot
panel(wui_firEUrisk, axes = FALSE)

```

```{r}
#| label: wui-firEUrisk-dist

wui_firEUrisk <- terra::readRDS("data/processed_data/wui_firEUrisk.rsd")

# Data by country
wui_firEUrisk_no <- mask(wui_firEUrisk, no_v)
wui_firEUrisk_se <- mask(wui_firEUrisk, se_v)

```

```{r}
#| label: tbl-wui-firEUrisk-dist
#| tbl-cap: Changes in WUI percentages of WUI by country (firEUrisk)
#| tbl-subcap: 
#|  - Norway
#|  - Sweden
#| layout-ncol: 2
#| cache: true

tbl_summary_distance(wui_firEUrisk_no)
tbl_summary_distance(wui_firEUrisk_se)

```

Aggregate all maps in one figure (Note: data are re-sampled to 500859 cells in the plots)
```{r}
#| label: fig-wui-vegetation-type-distance
#| fig-cap: WUI by vegetation type map and threshold distance
#| cache: true

# Plot
p_effis <- ggplot() + 
  geom_spatraster(data = wui_effis, maxcell = 2000000) +
  facet_wrap(~lyr, ncol = 1) +
  labs(title = "Effis") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))

p_corine <- ggplot() + 
  geom_spatraster(data = wui_corine, maxcell = 2000000) +
  facet_wrap(~lyr, ncol = 1) +
  labs(title = "Corine") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))

p_national <- ggplot() + 
  geom_spatraster(data = wui_national, maxcell = 2000000) +
  facet_wrap(~lyr, ncol = 1) +
  labs(title = "National") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))

p_firEUrisk <- ggplot() + 
  geom_spatraster(data = wui_firEUrisk, maxcell = 2000000) +
  facet_wrap(~lyr, ncol = 1) +
  labs(title = "firEUrisk") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))

p_effis + p_corine + p_national + p_firEUrisk +
  plot_layout(guides = "collect",
              ncol = 4)

```

Plot variation in the percentage of WUI with the distance to large forest areas by each vegetation/fuel map.

```{r}
#| label: fig-wui-distance
#| fig-cap: Change in the percentage of WUI with the theshold distance to large forest areas

get_tbl_pct <- function(.x) {

  t <- .x |>
    as_tibble() |> 
    drop_na() |> 
    pivot_longer(everything(),
                 names_to = "distance",
                 values_to = "value") |>
    group_by(distance, value) |>
    summarise(count = n()) |>
    ungroup() |> 
    group_by(distance) |>
    mutate(total = sum(count),
           pct = 100*count/total) |> 
    ungroup() |> 
    pivot_wider(names_from = "value",
                values_from = c("count", "pct")) |> 
    select(distance, starts_with("pct")) |> 
    mutate(distance = gsub("d", "", distance),
           distance = as.numeric(distance)) |> 
    arrange(distance)

  return(t)

}

# Combine rasters
raster_list <- list(wui_effis_no,
                    wui_corine_no,
                    wui_national_no,
                    wui_firEUrisk_no,
                    wui_effis_se,
                    wui_corine_se,
                    wui_national_se,
                    wui_firEUrisk_se)

# Tables with percentages
dist_list <- raster_list |> purrr::map(get_tbl_pct)

# Add names to de list
names(dist_list) <- c("Norway_effis",
                      "Norway_corine",
                      "Norway_national",
                      "Norway_firEUrisk",
                      "Sweden_effis",
                      "Sweden_corine",
                      "Sweden_national",
                      "Sweden_firEUrisk")

# Merge with names
dist_tbl <- bind_rows(dist_list, .id = "Ticker") |> 
  separate(Ticker, c("country", "vegetation"))

# add value for 2400 m 
get_tbl_pct_2 <- function(.x){

t <- .x |>
  as_tibble() |> 
  drop_na() |> 
  pivot_longer(everything(),
               names_to = "vegetation",
               values_to = "value") |>
  group_by(vegetation, value) |>
  summarise(count = n()) |>
  ungroup() |> 
  group_by(vegetation) |>
  mutate(total = sum(count),
         pct = 100*count/total) |> 
  ungroup() |> 
  pivot_wider(names_from = "value",
              values_from = c("count", "pct")) |> 
  select(vegetation, starts_with("pct"))

return(t)

}

# Plot
ggplot(data = dist_tbl,
       aes(x = distance,
           y = pct_Interface,
           col = vegetation)) +
  geom_point() +
  geom_line() +
  facet_grid(~country) +
  labs(x = "Distance [m]",
       y = "Perc. [%]"
       # title = "Percentage of WUI Interface vs. distance to forest"
       ) +
  theme_bw() +
  theme(strip.text.x = element_text(size = 12, face = "italic"))

```

### Effect threshold building density

In the following figure we can see the distribution of buildings density (#/$km^2$). The number of buildings in the buffer area is then calculated multiplying the building density by the area of the buffer (i.e., minimum number of buildings (Nmin) = density threshold \[#/$km^2$\] \* Buffer area \[$km^2$\]). In red is the standard threshold for classifying an area as possible WUI, then we have used half of this value (i.e., 3.08 #/$km^2$), representing approximately 25% of the data. Finally, we have also tested with 1 #/$km^2$, which is equivalent to classified any area with buildings as potential WUI.

```{r}
#| label: effect-threshold-build-density

build_focal_no <- mask(build_focal, no_v)

build_focal_se <- mask(build_focal, se_v)

# Get density table

get_d_counts <- function(.x) {

  build_focal |> 
    mask(.x) |> 
    as_tibble() |> 
    drop_na() |> 
    mutate(d_km2 = focal_sum / area_buffer)

}

df_counts <- list(no_v, se_v) |> 
  map(get_d_counts) 
names(df_counts) <- c("Norway", "Sweden")

df_counts_bind <- bind_rows(df_counts, .id = "country")

# Create tables with summary statistics to add to the plot
get_tbl_quantiles <- function(.x) {

  tbl <- .x |> 
    select(d_km2) |>
    quantile(prob = seq(0, 1, 0.25), na.rm = T) |>
    round(2) |> 
    as_tibble() |> 
    rename(d = value) |> 
    mutate(Quantile = c("Min", "Q1", "Median", "Q3", "Max"),
           Quantile = forcats::fct_reorder(Quantile, d)) |> 
    select(Quantile, d)

  tbl_quantiles <- tibble(x = 2800,
                          y = 4.0e+5, 
                          table = list(tbl))

  return(tbl_quantiles)

}

tbl_quantiles <- df_counts |>
  map(get_tbl_quantiles) |> 
  bind_rows(.id = "country")

# Plot
ggplot(df_counts_bind, aes(x = d_km2)) + 
  geom_histogram(bins = 50, color = "darkblue", fill = "lightblue") +
  facet_grid(~country) +
  scale_x_log10() +
  geom_vline(xintercept = 6.172,
             linetype = "dashed",
             color = "red",
             linewidth = 1.3) +
  annotate("text", x = 8, y = 420000,
           label = "6.172", colour = "red", parse = TRUE) + 
  geom_vline(xintercept = 3.08,
             linetype = "dashed",
             color = "orange",
             linewidth = 1.3) +
  annotate("text", x = 4, y = 420000,
           label = "3.08", colour = "orange", parse = TRUE) + 
  geom_vline(xintercept = 1.00,
             linetype = "dashed", 
             color = "blue",
             linewidth = 1.3) +
  annotate("text", x = 1.1, y = 420000,
           label = "1", colour = "blue", parse = TRUE) + 
  labs(title = "Building density",
       x = "#/km2",
       y = "") +
  theme_bw() +
  theme(legend.key = element_rect(fill = NA),
        legend.background = element_rect(color = "gray", linetype = "solid"),
        legend.position = c(.95, .95),
        legend.justification = c("right", "top"),
        legend.box.just = "right",
        legend.margin = margin(6, 6, 6, 6),
        strip.text.x = element_text(size = 12, face = "italic")) +
  # Add tabble with summary statistics
  ggpp::geom_table(data = tbl_quantiles,
                   aes(x = x, y = y, label = table),
                   table.theme = ggpp::ttheme_gtplain(base_size = 12))

```

Potential WUI based on different building density:

#### Building density 3.08 #/km2

```{r}

# Minimum number of buildings (N) in the buffer area ----
# WUI if building density (d) > 6.17 [#/km^2]
# N = d [#/km^2] * A [km^2]
# A = pi * radio^2

N_min_3 <- round(3.08 * area_buffer, 3)

wui_potential_3 <- build_focal
wui_potential_3[wui_potential_3 <= N_min_3] <- NA
wui_potential_3[wui_potential_3 >  N_min_3] <- 1

plot(region_r, col = "#E7EEE7", axes = F, legend = F)
plot(wui_potential_3, axes = F, legend = F, add = T)

```

```{r}
#| label: calculate-wui3-region
#| cache: true

# EFFIS
effis_wui_3 <- get_wui_raster(.veg_perc = effis_wv_perc,
                              .veg_dist = effis_wild_veg_large_dist_2400,
                              .pot_wui = wui_potential_3)

# CORINE
corine_wui_3 <- get_wui_raster(.veg_perc = corine_wv_perc,
                               .veg_dist = corine_wild_veg_large_dist_2400,
                               .pot_wui = wui_potential_3)

# National maps
sr16_wui_3 <- get_wui_raster(.veg_perc = sr16_wv_perc,
                             .veg_dist = sr16_wild_veg_large_dist_2400,
                             .pot_wui = wui_potential_3,
                             .region = no_v)

nrm_wui_3 <- get_wui_raster(.veg_perc = nmd_wv_perc,
                            .veg_dist = nmd_wild_veg_large_dist_2400,
                            .pot_wui = wui_potential_3,
                            .region = se_v)

national_wui_3 <- mosaic(sr16_wui_3, nrm_wui_3, fun = "max") |> 
  add_colours_wui()

# firEUrisk
firEUrisk_wui_3 <- get_wui_raster(.veg_perc = firEUrisk_wv_perc,
                                  .veg_dist = firEUrisk_wild_veg_large_dist_2400,
                                  .pot_wui = wui_potential_3)

# Put all rasters in one 
wui_r_3 <- c(effis_wui_3, corine_wui_3, national_wui_3, firEUrisk_wui_3)
names(wui_r_3) <- c("effis", "corine", "national", "firEUrisk")

terra::saveRDS(wui_r_3, "data/processed_data/wui_r_3.rds")

```

Map of WUI areas (@fig-wui3-maps-region; @tbl-pct-wui3-region).

```{r}
#| label: read-wui3-raster

wui_r_3 <- terra::readRDS("data/processed_data/wui_r_3.rds")

```

```{r}
#| label: fig-wui3-maps-region
#| fig-cap: "WUI maps of Norway based on different fuel/vegetation maps (resolution of 250 x 250 m)"
#| fig-subcap: 
#|   - "EFFIS"
#|   - "CORINE"
#|   - "FirEUrisk"
#|   - "National databases"
#| layout-ncol: 2
#| fig-width: 4
#| fig-format: png
#| cache: true

f_plot <- function(.x){ plot(.x, 
                             axes = F, 
                             mar = c(0, 0, 0, 0),
                             legend = "topleft")  }

f_plot(wui_r_3$effis)
f_plot(wui_r_3$corine)
f_plot(wui_r_3$firEUrisk)
f_plot(wui_r_3$national)

```

```{r}
#| label: tbl-pct-wui3-region
#| tbl-cap: Percentages of WUI by vegetation
#| tbl-subcap: 
#|  - Norway
#|  - Sweden
#| layout-ncol: 2

# Summary WUI by country
wui_r_3_no <- mask(wui_r_3, no_v)
wui_r_3_se <- mask(wui_r_3, se_v)

# summary table 
tbl_summary_vegetation(wui_r_3_no)
tbl_summary_vegetation(wui_r_3_se)

```

#### Building density 1 #/km2

Equivalent to select any area with building as WUI.

```{r}

# Minimum number of buildings (N) in the buffer area ----
# WUI if building density (d) > 6.17 [#/km^2]
# N = d [#/km^2] * A [km^2]
# A = pi * radio^2

N_min_1 <- round(1.00 * area_buffer, 3)

wui_potential_1 <- build_focal
wui_potential_1[wui_potential_1 <= N_min_1] <- NA
wui_potential_1[wui_potential_1 >  N_min_1] <- 1

plot(region_r, col = "#E7EEE7", axes = F, legend = F)
plot(wui_potential_1, axes = F, legend = F, add = T)

```

```{r}
#| label: calculate-wui1-region
#| cache: true

# EFFIS
effis_wui_1 <- get_wui_raster(.veg_perc = effis_wv_perc,
                              .veg_dist = effis_wild_veg_large_dist_2400,
                              .pot_wui = wui_potential_1)

# CORINE
corine_wui_1 <- get_wui_raster(.veg_perc = corine_wv_perc,
                               .veg_dist = corine_wild_veg_large_dist_2400,
                               .pot_wui = wui_potential_1)

# National maps
sr16_wui_1 <- get_wui_raster(.veg_perc = sr16_wv_perc,
                             .veg_dist = sr16_wild_veg_large_dist_2400,
                             .pot_wui = wui_potential_1,
                             .region = no_v)

nrm_wui_1 <- get_wui_raster(.veg_perc = nmd_wv_perc,
                            .veg_dist = nmd_wild_veg_large_dist_2400,
                            .pot_wui = wui_potential_1,
                            .region = se_v)

national_wui_1 <- mosaic(sr16_wui_1, nrm_wui_1, fun = "max") |> 
  add_colours_wui()

# firEUrisk
firEUrisk_wui_1 <- get_wui_raster(.veg_perc = firEUrisk_wv_perc,
                                  .veg_dist = firEUrisk_wild_veg_large_dist_2400,
                                  .pot_wui = wui_potential_1)

# Put all rasters in one 
wui_r_1 <- c(effis_wui_1, corine_wui_1, national_wui_1, firEUrisk_wui_1)
names(wui_r_1) <- c("effis", "corine", "national", "firEUrisk")

terra::saveRDS(wui_r_1, "data/processed_data/wui_r_1.rds")

```

Map of WUI areas (@fig-wui1-maps-region; @tbl-pct-wui1-region).

```{r}
#| label: read-wui1-raster

wui_r_1 <- terra::readRDS("data/processed_data/wui_r_1.rds")

```

```{r}
#| label: fig-wui1-maps-region
#| fig-cap: "WUI maps of Norway based on different fuel/vegetation maps (resolution of 250 x 250 m)"
#| fig-subcap: 
#|   - "EFFIS"
#|   - "CORINE"
#|   - "FirEUrisk"
#|   - "National databases"
#| layout-ncol: 2
#| fig-width: 4
#| fig-format: png
#| cache: true

f_plot <- function(.x){ plot(.x, 
                             axes = F, 
                             mar = c(0, 0, 0, 0),
                             legend = "topleft")  }

f_plot(wui_r_1$effis)
f_plot(wui_r_1$corine)
f_plot(wui_r_1$firEUrisk)
f_plot(wui_r_1$national)

```

```{r}
#| label: tbl-pct-wui1-region
#| tbl-cap: Percentages of WUI by vegetation
#| tbl-subcap: 
#|  - Norway
#|  - Sweden
#| layout-ncol: 2

# Summary WUI by country
wui_r_1_no <- mask(wui_r_1, no_v)
wui_r_1_se <- mask(wui_r_1, se_v)

# Tables
tbl_summary_vegetation(wui_r_1_no)
tbl_summary_vegetation(wui_r_1_se)

```

Plot all together
```{r}
#| label: fig-wui-threshold

p_1 <- ggplot() + 
  geom_spatraster(data = wui_r, maxcell = 2000000) +
  facet_wrap(~lyr, nrow = 1) +
  labs(title = "d = 1.00 #/km2") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))

p_3.08 <- ggplot() + 
  geom_spatraster(data = wui_r, maxcell = 2000000) +
  facet_wrap(~lyr, nrow = 1) +
  labs(title = "d = 3.08 #/km2") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))

p_6.17 <- ggplot() + 
  geom_spatraster(data = wui_r, maxcell = 2000000) +
  facet_wrap(~lyr, nrow = 1) +
  labs(title = "d = 6.17 #/km2") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))

p_1 / p_3.08 / p_6.17 +
  plot_layout(guides = "collect")

df_col <- data.frame(id = 0:2, col = c("#000000", "#E69F00", "#009E73"))
df_col_lst <- replicate(nlyr(wui_r), df_col, simplify = FALSE)
coltab(wui_r) <- df_col_lst
plot(wui_r)

```

# Wildfires datasets

We analyse all wildfires (i.e., 330) that have affected buildings from 2016 to 2022. The data have been obtained from XX (Norway) and XXX (Sweden).

```{r}
#| label: wildfires-points

wf_no <- readxl::read_xlsx("data/wui_fires_norway_wgs84.xlsx", skip = 1) |> 
  janitor::clean_names() |> 
  st_as_sf(coords = c("geolokasjon_lengdegrad", "geolokasjon_breddegrad"),
           crs =  4326) |> 
  select(geometry) |> 
  st_transform(crs = 3035) |> 
  mutate(country = "NO")

wf_se <- readxl::read_xlsx("data/wui_fires_sweden_epsg3006.xlsx") |> 
  janitor::clean_names() |> 
  mutate(datetime = lubridate::as_date(datetime)) |> 
  filter(datetime >= lubridate::as_date("2016-01-01")) |> 
  st_as_sf(coords = c("ecoord", "ncoord"),
           crs =  3006) |> 
  select(geometry) |> 
  st_transform(crs = 3035) |> 
  mutate(country = "SE")

wf <- rbind(wf_no, wf_se)

ggplot() +
  geom_sf(data = region_sf, fill = "#E7EEE7") +
  geom_sf(data = country_sf, fill = NA) +
  geom_sf(data = wf, col = "red", size = 0.4) +
  theme_void()

```

Intersect points with WUI maps

## Effect distance to large forest

Get if the grid to which the wildfire is located was defined as WUI in the respective maps.

```{r}
#| label: tbl-wildfire-wui-map
#| tbl-cap: "Percentage of wildfires in each WUI class"

# Detect wui class for each point (wildfire)
extract_wui <- function(.x) { extract(.x, wf) }

wf_wui <- list(wui_effis, wui_corine, wui_national, wui_firEUrisk) |> 
  map(extract_wui) |> 
  map(as_tibble)
names(wf_wui) <- c("effis", "corine", "national", "firEUrisk")

# Tables with counts of wildfire in each WUI class
f_counts <- function(.x) {
  .x |> 
  pivot_longer(-ID) |> 
  count(name, value) |> 
  mutate(name = factor(name, 
                       levels = c("d200",
                                  "d500",
                                  "d1000",
                                  "d2000",
                                  "d2400"),
                       ordered = T)) |>
  arrange(name) |> 
  pivot_wider(names_from = name,
              values_from = n) |> 
  mutate(across(.cols = starts_with("d"),
                .fns = ~ 100 * (.x / sum(.x)),
                .names = "perc_{.col}"))

}

wf_wui_counts <- wf_wui |>
  map(f_counts) |> 
  # Merge with names
  bind_rows(.id = "vegetation")

# Print table
wf_wui_counts |> 
  # select("vegetation", "value", starts_with("perc")) |> 
  gt(groupname_col = "vegetation",
     row_group_as_column = FALSE) |> 
  fmt_number(columns = starts_with("perc"),
             decimals = 1) |> 
  cols_label(
    vegetation = "",
    value = md("**WUI class**"),
    perc_d200 = md("**d200**"),
    perc_d500 = md("**d500**"),
    perc_d1000 = md("**d1000**"),
    perc_d2000 = md("**d2000**"),
    perc_d2400 = md("**d2400**")
  )  |> 
  cols_label_with(fn = function(x) {
    janitor::make_clean_names(x, case = "title") |>
      tolower() |>
      stringr::str_replace_all("^|$", "**") |>
      md()
  }) |> 
  tab_spanner(
    label = md("**Percentage**"),
    columns = starts_with("perc")
  ) |> 
  tab_spanner(
    label = md("**Count**"),
    columns = starts_with("d")
  ) |> 
  tab_options(
    table.font.size = 9,
    row_group.background.color = "#E7EEE7",
    row_group.padding = px(1),
    row_group.font.weight = "bold",
    data_row.padding = px(1)
  )

```

## Effect threshold building density

```{r}
#| label: read-wui-maps-building-density

wui_r_1 <- terra::readRDS("data/processed_data/wui_r_1.rds")
wui_r_3 <- terra::readRDS("data/processed_data/wui_r_3.rds")
wui_r <- terra::readRDS("data/processed_data/wui_r.rds")

```

```{r}
#| label: tbl-wildfire-wui-map-building-density
#| tbl-cap: "Percentage of wildfires in WUI calss based on building density threshold"

# Detect wui class for each point (wildfire)
extract_wui <- function(.x) { extract(.x, wf) }

wf_wui_bd <- list(wui_r_1, wui_r_3, wui_r) |> 
  map(extract_wui) |> 
  map(as_tibble)
names(wf_wui_bd) <- c("bd_1", "bd_3", "bd_6")

# Tables with counts og the N. wildfire in each wui class
f_counts <- function(.x) {
  .x |> 
  pivot_longer(-ID) |> 
  count(name, value) |> 
  mutate(name = factor(name, 
                       levels = c("effis",
                                  "corine",
                                  "national",
                                  "firEUrisk"),
                       ordered = T)) |>
  arrange(name) |> 
  pivot_wider(names_from = name,
              values_from = n) |> 
  mutate(across(.cols = -value,
                .fns = ~ 100 * (.x / sum(.x)),
                .names = "perc_{.col}"))

}

wf_wui_bd_counts <- wf_wui_bd |>
  map(f_counts) |> 
  # Merge with names
  bind_rows(.id = "threshold")

# Print table
wf_wui_bd_counts |> 
  mutate(definition = case_when(
    threshold == "bd_1" ~ "threshold = 1.00 #/km2",
    threshold == "bd_3" ~ "threshold = 3.08 #/km2",
    threshold == "bd_6" ~ "threshold = 6.17 #/km2"
  )) |> 
  select(-threshold) |> 
  gt(groupname_col = "definition",
     row_group_as_column = FALSE) |> 
  fmt_number(columns = starts_with("perc"),
             decimals = 1) |> 
  cols_label(
    definition = "",
    value = md("**WUI class**"),
    effis = md("**effis**"),
    corine = md("**corine**"),
    national = md("**national**"),
    firEUrisk = md("**firEUrisk**"),
    perc_effis = md("**effis**"),
    perc_corine = md("**corine**"),
    perc_national = md("**national**"),
    perc_firEUrisk = md("**firEUrisk**")
  )  |> 
  tab_spanner(
    label = md("**Percentage**"), columns = starts_with("perc")
  ) |> 
  tab_spanner(
    label = md("**Count**"), columns = c(effis, corine, national, firEUrisk)) |> 
  tab_options(
    table.font.size = 9,
    row_group.background.color = "#E7EEE7",
    row_group.padding = px(1),
    row_group.font.weight = "bold",
    data_row.padding = px(1)
  )

```

### 2024-08-31 article.qmd
Add charts of WUI percentage with geofacet

# Aricle

How to approach the definition of WUI in northern Europe (doi: )

## Abstract {.unnumbered}

This paper aims to delineate the wildland-urban interface in the sparsely populated, limited resourced Atlantic-Boreal Europe. Due to its demographic and biophysical differences to southern Europe, as well as to other fire-prone regions, current WUI definitions to suit the needs of southern countries are not viable. Therefore, common WUI mapping assumptions, and how well the resulting WUI maps (thresholds) capture the reality of wildfire in northern Europe, are scoped.

## Acknowledgements {.unnumbered}

This work has been financed by XXX

### 2024-08-31 article_codes.qmd
Add charts of WUI percentage with geofacet

--- 2024-08-19 raster.qmd
@@
 
-# National WUI map
+# R scripts
 
@@
 #' @param .threshold Proportion: from 0-1 (default  = 0.5; i.e., 50%) 
-#' @param .cache_path File path to save the results 
 get_perc_veg_buffer <- function(.veg,
@@
 #' @param .x Raster with distances (from `terra::distance`) 
-#' @param .max_distance Maximun distance in m (e.g., 2400 m)
+#' @param .max_distance Maximum distance in [m] (e.g., 2400 m)
 get_dist_max <- function(.x, .max_dist) {
@@
   coltab(wui) <- data.frame(id = 0:2, 
-                            col = c("#E7EEE7", "#FF6633", "#003366"))
+                            col = c("#FAFAFA", "#FF6633", "#003366"))
 
@@
   # Add colour table 
-  df_col <- data.frame(id = 0:2, col = c("#E7EEE7", "#FF6633", "#003366"))
+  df_col <- data.frame(id = 0:2, col = c("lightgrey", "#FF6633", "#003366"))
   df_col_lst <- replicate(nlyr(r), df_col, simplify = FALSE)
@@
 # Create local directory for caching big files
+dir.create("C:/GISCO_spatial_data")
 options(gisco_cache_dir = "C:/GISCO_spatial_data")
@@
 
+dir.create("data/processed_data")
 dir.create("data/processed_data/tiles")
@@
   theme_bw() +
-  theme(panel.background = element_rect(fill = "#E7EEE7")) +
+  theme(panel.background = element_rect(fill = "lightgrey")) +
   theme(panel.grid = element_blank(),
@@
 
-plot(region_r, col = "#E7EEE7", axes = FALSE, legend = FALSE)
+plot(region_r, col = "lightgrey", axes = FALSE, legend = FALSE)
 plot(effis_fuel_cls, 
@@
 # Wildand vegetation (Yes = 1, No = NA)
-plot(region_r, col = "#E7EEE7", axes = F, legend = F)
+plot(region_r, col = "lightgrey", axes = F, legend = F)
 plot(effis_wild_veg, col = "#009E73", axes = F, legend = F, add = T)
@@
 # Within a distance of 2.4 km (Yes = 1, No = 0)
-plot(region_r, col = "#E7EEE7", axes = F, legend = F)
+plot(region_r, col = "lightgrey", axes = F, legend = F)
 plot(effis_wild_veg_large_dist_2400, col = c(NA, "#0072B2"), axes = F, legend = F, add = T)
@@
 # Plot ----
-plot(region_r, col = "#E7EEE7", axes = FALSE, legend = FALSE)
+plot(region_r, col = "lightgrey", axes = FALSE, legend = FALSE)
 plot(corine_fuel_veg, col = clc2018_wild_veg_list$rgb_code, 
@@
 # Wildland vegetation (Yes = 1, No = NA)
-plot(region_r, col = "#E7EEE7", axes = F, legend = F)
+plot(region_r, col = "lightgrey", axes = F, legend = F)
 plot(corine_wild_veg, col = "#009E73", axes = F, legend = F, add = T)
@@
 # Within a distance of 2.4 km (Yes = 1, No = 0)
-plot(region_r, col = "#E7EEE7", axes = F, legend = F)
+plot(region_r, col = "lightgrey", axes = F, legend = F)
 plot(corine_wild_veg_large_dist_2400, col = c(NA, "#0072B2"), axes = F, legend = F, add = T)
@@
 no_v_p <- project(no_v, "EPSG:25833")
-plot(no_v_p, col = "#E7EEE7", axes = FALSE, legend = FALSE)
+plot(no_v_p, col = "lightgrey", axes = FALSE, legend = FALSE)
 plot(sr16_vrt, col = c("lightgreen", "darkgreen", "yellow"), axes = FALSE, add = T)
@@
 # Wildland vegetation (Yes = 1, No = NA)
-plot(region_r, col = "#E7EEE7", axes = F, legend = F)
+plot(region_r, col = "lightgrey", axes = F, legend = F)
 plot(sr16_wild_veg, col = "#009E73", axes = F, legend = F, add = T)
@@
 # Within a distance of 2.4 km (Yes = 1, No = 0)
-plot(region_r, col = "#E7EEE7", axes = F, legend = F)
+plot(region_r, col = "lightgrey", axes = F, legend = F)
 plot(sr16_wild_veg_large_dist_2400, 
@@
 
-plot(region_r, col = "#E7EEE7", axes = FALSE, legend = FALSE)
+plot(region_r, col = "lightgrey", axes = FALSE, legend = FALSE)
 plot(firEUrisk_fuel_250m, axes = FALSE, add = T)
@@
 # Wildand vegetation (Yes = 1, No = NA)
-plot(region_r, col = "#E7EEE7", axes = F, legend = F)
+plot(region_r, col = "lightgrey", axes = F, legend = F)
 plot(firEUrisk_wild_veg, col = "#009E73", axes = F, legend = F, add = T)
@@
 # Within a distance of 2.4 km (Yes = 1, No = 0)
-plot(region_r, col = "#E7EEE7", axes = F, legend = F)
+plot(region_r, col = "lightgrey", axes = F, legend = F)
 plot(firEUrisk_wild_veg_large_dist_2400, col = c(NA, "#0072B2"), axes = F, legend = F, add = T)
@@
 flowchart TD
-  A[Grid 250m x 250m] --> B{"Building density > 6.17 #/km2"}
-  B --  Yes  --> C{Vegetation > 50%}
-  B --  No   --> D[No WUI]
-  C --  Yes  --> E[Intermix] 
-  C --  No   --> F{within 2.4 km of large forests (>5 km2)}
-  F --  Yes  --> G[Interface]
-  F --  No   --> H[No WUI]
-  style E fill:#CC79A7
-  style G fill:#F0E442
-  style D fill:#D3D3D3
-  style H fill:#D3D3D3
+    A["Grid 250m x 250m"] --> B{"Building density > 6.17 #/km2"}
+    B -- Yes --> C{"Vegetation > 50%"}
+    B -- No --> D["No WUI"]
+    C -- Yes --> E["Intermix"]
+    C -- No --> F{"Within 2.4 km of large 
+    forest (>5 km)"}
+    F -- Yes --> G["Interface"]
+    F -- No --> H["No WUI"]
+    style D fill:lightgrey
+    style E fill:#003366,color:#FFFFFF
+    style G fill:#FF6633
+    style H fill:lightgrey
+
 ```
@@
 
-3.  Create a raster indicating if the grid cell is within 2.4 km of a large (>5 km2) fuel vegetation area (`wild_veg_large_dist_*`)
+3.  Create a raster indicating if the grid cell is within 2.4 km of a large (\>5 km2) fuel vegetation area (`wild_veg_large_dist_*`)
 
@@
 
-plot(region_r, col = "#E7EEE7", axes = F, legend = F)
+plot(region_r, col = "lightgrey", axes = F, legend = F)
 plot(wui_potential, axes = F, legend = F, add = T)
@@
 #     data.frame(id = 0:2, 
-#                col = c("#E7EEE7", "#F0E442", "#CC79A7"))
+#                col = c("lightgrey", "#F0E442", "#CC79A7"))
 #   
@@
       table.font.size = 9,
-      row_group.background.color = "#E7EEE7",
+      row_group.background.color = "lightgrey",
       row_group.padding = px(1),
@@
 ```
-
-Area of wild vegetation in each county: Number vegetation cells (centroid is inside the polygon), sum then and calculate the area based on the raster resolution
+```{r}
+#| label: fig-pct-wui-counties
+
+# Add county ID
+counties_cent_ID <- counties |>
+  st_centroid() %>%
+  mutate(x = st_coordinates(.)[,1],
+         y = st_coordinates(.)[,2]) |> 
+  arrange(desc(y), x) |> 
+  mutate(ID = seq_along(counties$NUTS_NAME))
+
+counties_ID <- counties |> 
+  as_tibble() |> 
+  left_join(counties_cent_ID, "NUTS_NAME") |> 
+  st_sf()
+
+# geofacet
+
+mygrid <- data.frame(
+  code = c("1 Troms og F.", 
+           "2 Norrbotten", 
+           "3 Nordland", 
+           "4 Västerbotten", 
+           "5 Trøndelag", 
+           "7 Västernorrland", 
+           "6 Jämtland", 
+           "8 Møre og R.", 
+           "10 Innlandet", 
+           "11 Dalarna", 
+           "9 Gävleborg", 
+           "12 Vestland",
+           "15 Oslo",
+           "14 Viken",
+           "13 Uppsala",
+           "16 Västmanland",
+           "20 Örebro", 
+           "17 Värmland",
+           "22 Rogaland", 
+           "19 Vestfold og T.",
+           "18 Stockholm", 
+           "21 Södermanland",
+           "24 Östergötland",
+           "25 Västra Götaland",
+           "23 Agder", 
+           "29 Halland", 
+           "27 Jönköping", 
+           "28 Kalmar", 
+           "26 Gotland", 
+           "30 Kronoberg", 
+           "31 Blekinge",
+           "32 Skåne"),
+  row = c(2, 2, 3, 3, 4, 4, 4, 5, 5, 5,
+          5, 6, 6, 6, 6, 6, 6, 6, 7, 7, 
+          7, 7, 7, 7, 8, 8, 8, 8, 9, 9, 9, 10),
+  col = c(6, 7, 6, 7, 5, 7, 6, 4, 5, 6,
+          7, 3, 5, 4, 9, 8, 7, 6, 3, 4, 
+          9, 8, 7, 6, 4, 6, 7, 8, 9, 6, 7, 6),
+  name = c("Troms og Finnmark",
+           "Norrbotten", 
+           "Nordland",
+           "Västerbotten", 
+           "Trøndelag", 
+           "Västernorrland",
+           "Jämtland", 
+           "Møre og Romsdal",
+           "Innlandet", 
+           "Dalarna", 
+           "Gävleborg",
+           "Vestland", 
+           "Oslo", 
+           "Viken", 
+           "Uppsala",
+           "Västmanland",
+           "Örebro", 
+           "Värmland",
+           "Rogaland", 
+           "Vestfold og Telemark",
+           "Stockholm", 
+           "Södermanland", 
+           "Östergötland", 
+           "Västra Götaland", 
+           "Agder", 
+           "Halland", 
+           "Jönköping", 
+           "Kalmar",
+           "Gotland", 
+           "Kronoberg", 
+           "Blekinge", 
+           "Skåne"),
+  stringsAsFactors = FALSE
+)
+
+# Maps
+
+p_map <- ggplot() +
+  geom_sf(data = counties_ID, 
+          fill = "#FAFAFA",
+          col = "lightgrey",
+          size = 0.5) +
+  geom_sf_text(data = counties_ID, 
+               aes(label = ID),
+               col = "red",
+               size = 3) +
+  theme_void()
+
+p_bar <- counties_wui |> 
+  select(NUTS_NAME, name, starts_with("pct")) |> 
+  select(-pct_wui) |> 
+  pivot_longer(cols = starts_with("pct"), 
+               names_to = "type",
+               values_to = "value") |> 
+  mutate(type = str_remove_all(type, "pct_"),
+         NUTS_NAME = str_remove_all(NUTS_NAME, "s län"),
+         NUTS_NAME = str_remove_all(NUTS_NAME, " län")) |> 
+  mutate(name = case_when(name == "corine" ~ "C",
+                          name == "effis" ~ "E",
+                          name == "national" ~ "N",
+                          name == "firEUrisk" ~ "F")) |>
+  mutate(name = factor(name, levels = c("N", "C", "E", "F"))) |> 
+  mutate(type = factor(type, levels = c("No_wui", "Interface", "Intermix"))) |> 
+  ggplot() +
+  geom_bar(aes(x = name,
+               y = value,
+               fill = type),
+           col = "grey",
+           stat = "identity",
+           position = "fill"
+           ) +
+  scale_fill_manual(values = c("#FAFAFA", "#FF6633", "#003366")) +
+  geofacet::facet_geo(~NUTS_NAME,
+                      grid = mygrid, 
+                      label = "code"
+                      ) +
+  labs(x = "",
+       y = "") +
+  theme_bw() +
+  theme(
+    strip.text = element_text(size = 9, face = "bold")
+  )
+
+ggsave(plot = p_map, 
+       filename = "figures/counties_map.png", 
+       width = 3, height = 3)
+
+ggsave(plot = p_bar, 
+       filename = "figures/counties_wui_perc.png", 
+       width = 15, height = 20)
+
+```
+
+![](figures/counties_map.png)
+![WUI percentage by county](figures/counties_wui_perc.png)
+
+Area of wild vegetation by county (@tbl-summary-table-areas): Number vegetation cells (centroid is inside the polygon), sum then and calculate the area based on the raster resolution.
 
@@
 #| label: tbl-summary-table-areas
-#| tbl-cap: Number of buildings and wildland vegetation percentage by county 
+#| tbl-cap: Number of buildings and wildland vegetation percentage by county (and total)
 #| cache: true
 
-# Number of buildings (use building_count?) ----
+# Number of buildings by county  ----
 county_build_count <- build_counts |> 
@@
 
-# Percentage vegetation county ----
-# EFFIS
-effis_county <- effis_wild_veg |> 
+# Percentage vegetation by county ----
+
+# Function for getting vegetation area by counties 
+f_pct_veg <- function(df) {
+  df |> # sf object with vegetation data  
     # Number of cells
@@
     mutate(
-    area_km = wild_veg * ((res(effis_wild_veg))[1]/1000 * (res(effis_wild_veg ))[2]/1000)
+      area_km = wild_veg * ((res(df))[1]/1000 * (res(df ))[2]/1000)
     ) |> 
@@
     st_drop_geometry() |> 
-  select(NUTS_NAME, CNTR_CODE, area_km) |> 
+    select(NUTS_NAME, CNTR_CODE, area_km)
+}
+
+## Run for each vegetation type
+
+# Effis
+effis_county  <- f_pct_veg(effis_wild_veg) |> 
     rename(area_km_effis = area_km)
 
-# CORINE
-corine_county <- corine_wild_veg |> 
-  # Number of cells
-  extract(y = vect(counties),
-          fun = "sum",
-          na.rm = TRUE,
-          bind = TRUE) |>  
-  st_as_sf() |> 
-  # Area (number cells x area cells)
-  mutate(
-    area_km = wild_veg * ((res(corine_wild_veg))[1]/1000 * (res(corine_wild_veg ))[2]/1000)
-    ) |> 
-  # Get only values 
-  st_drop_geometry() |> 
-  select(NUTS_NAME, CNTR_CODE, area_km) |> 
+# Corine
+corine_county <- f_pct_veg(corine_wild_veg) |> 
   rename(area_km_corine = area_km)
@@
 # National
-sr16_county <- sr16_wild_veg |> 
-  # Number of cells
-  extract(y = vect(counties),
-          fun = "sum",
-          na.rm = TRUE,
-          bind = TRUE) |>  
-  st_as_sf() |> 
-  # Area (number cells x area cells)
-  mutate(
-    area_km = wild_veg * ((res(sr16_wild_veg))[1]/1000 * (res(sr16_wild_veg ))[2]/1000)
-    ) |> 
-  # Get only values 
-  st_drop_geometry() |> 
-  select(NUTS_NAME, CNTR_CODE, area_km) |> 
+sr16_county <- f_pct_veg(sr16_wild_veg) |> 
   filter(CNTR_CODE  == "NO")
-
-nmd_county <- nmd_wild_veg |> 
-  # Number of cells
-  extract(y = vect(counties),
-          fun = "sum",
-          na.rm = TRUE,
-          bind = TRUE) |>  
-  st_as_sf() |> 
-  # Area (number cells x area cells)
-  mutate(
-    area_km = wild_veg * ((res(nmd_wild_veg))[1]/1000 * (res(nmd_wild_veg ))[2]/1000)
-         ) |> 
-  # Get only values 
-  st_drop_geometry() |> 
-  select(NUTS_NAME, CNTR_CODE, area_km) |> 
+nmd_county  <- f_pct_veg(nmd_wild_veg)  |> 
   filter(CNTR_CODE  == "SE")
@@
 # FirEUrisk
-firEUrisk_county <- firEUrisk_wild_veg |> 
-  # Number of cells
-  extract(y = vect(counties),
-          fun = "sum",
-          na.rm = TRUE,
-          bind = TRUE) |>  
-  st_as_sf() |> 
-  # Area (number cells x area cells)
-  mutate(
-    area_km = wild_veg * ((res(firEUrisk_wild_veg))[1]/1000 * (res(firEUrisk_wild_veg ))[2]/1000)
-         ) |> 
-  # Get only values 
-  st_drop_geometry() |> 
-  select(NUTS_NAME, CNTR_CODE, area_km) |> 
+firEUrisk_county <- f_pct_veg(firEUrisk_wild_veg) |> 
   rename(area_km_firEUrisk = area_km)
@@
 # Joint tables
-county_veg <- effis_county |> 
+county_veg <- national_county |> 
   left_join(corine_county) |> 
-  left_join(national_county) |> 
+  left_join(effis_county) |> 
   left_join(firEUrisk_county)
@@
 
-# print table
-df_summary |> 
+# Summary tables by country and county ----
+summary_country <- df_summary |> 
+  select(CNTR_CODE, sum, area, starts_with("area_km")) |> 
+  group_by(CNTR_CODE) |> 
+  summarise_all(sum) |> 
+  mutate(NUTS_NAME = "Total",
+         build_den = sum/area) |> 
+  mutate(across(.cols = starts_with("area_km"),
+                .fns = ~ 100 * (.x / as.numeric(area)),
+                .names = "perc_{.col}")) |> 
+  relocate(build_den, .after = area) |> 
+  relocate(NUTS_NAME, .after = CNTR_CODE) |> 
   select(CNTR_CODE, NUTS_NAME, area,
@@
          CNTR_CODE = case_when(CNTR_CODE == "NO" ~ "NORWAY",
-                               CNTR_CODE == "SE" ~ "SWEDEN")) |> 
+                               CNTR_CODE == "SE" ~ "SWEDEN"))
+
+summary_county <- df_summary |> 
+  select(CNTR_CODE, NUTS_NAME, area,
+         sum, build_den,
+         starts_with("perc")) |>
+  arrange(CNTR_CODE, NUTS_NAME) |> 
+  mutate(area = as.numeric(area),
+         build_den = as.numeric(build_den),
+         CNTR_CODE = case_when(CNTR_CODE == "NO" ~ "NORWAY",
+                               CNTR_CODE == "SE" ~ "SWEDEN"))
+
+# Print table ----
+summary_county |> 
+  rbind(summary_country) |> 
   gt(groupname_col = "CNTR_CODE",
@@
       build_den = md("**Density [#/km2]**"),
+      perc_area_km_national = md("**National**"),
       perc_area_km_effis = md("**Effis**"),
       perc_area_km_corine = md("**Corine**"),
-      perc_area_km_national = md("**National**"),
       perc_area_km_firEUrisk = md("**FirEurisk**")
+    ) |> 
+  tab_style(
+    style = cell_text(weight = "bold"),  # Apply bold text
+    locations = cells_body(
+      rows = c(33, 34)  # specify the row number
+    )
   )
@@
 Aggregate all maps in one figure (Note: data are re-sampled to 500859 cells in the plots)
+
 ```{r}
@@
 #| fig-cap: WUI by vegetation type map and threshold distance
+#| fig-height: 15
 #| cache: true
@@
 
-plot(region_r, col = "#E7EEE7", axes = F, legend = F)
+plot(region_r, col = "lightgrey", axes = F, legend = F)
 plot(wui_potential_3, axes = F, legend = F, add = T)
@@
 
-plot(region_r, col = "#E7EEE7", axes = F, legend = F)
+plot(region_r, col = "lightgrey", axes = F, legend = F)
 plot(wui_potential_1, axes = F, legend = F, add = T)
@@
 Plot all together
+
 ```{r}
 #| label: fig-wui-threshold
+#| fig-height: 15
 
 p_1 <- ggplot() + 
-  geom_spatraster(data = wui_r, maxcell = 2000000) +
+  geom_spatraster(data = wui_r) +
   facet_wrap(~lyr, nrow = 1) +
@@
 p_3.08 <- ggplot() + 
-  geom_spatraster(data = wui_r, maxcell = 2000000) +
+  geom_spatraster(data = wui_r) +
   facet_wrap(~lyr, nrow = 1) +
@@
 p_6.17 <- ggplot() + 
-  geom_spatraster(data = wui_r, maxcell = 2000000) +
+  geom_spatraster(data = wui_r) +
   facet_wrap(~lyr, nrow = 1) +
@@
 
+# NOTE: geom_spatraster change the resolution of the map
 p_1 / p_3.08 / p_6.17 +
@@
 
-df_col <- data.frame(id = 0:2, col = c("#000000", "#E69F00", "#009E73"))
-df_col_lst <- replicate(nlyr(wui_r), df_col, simplify = FALSE)
-coltab(wui_r) <- df_col_lst
-plot(wui_r)
-
 ```
@@
 
-We analyse all wildfires (i.e., 330) that have affected buildings from 2016 to 2022. The data have been obtained from XX (Norway) and XXX (Sweden).
+We analyse all wildfires that have affected buildings from 2016 to 2022 (i.e., 330). The data have been obtained from XX (Norway) and XXX (Sweden).
 
@@
 ggplot() +
-  geom_sf(data = region_sf, fill = "#E7EEE7") +
+  geom_sf(data = region_sf, fill = "lightgrey") +
   geom_sf(data = country_sf, fill = NA) +
@@
     table.font.size = 9,
-    row_group.background.color = "#E7EEE7",
+    row_group.background.color = "lightgrey",
     row_group.padding = px(1),

### 2024-09-11 article_codes.qmd
Change wilfire data (now from 2016 to 2023)

@@
 
-# Function: get_perc_veg_buffer ----
+#' Function: get_perc_veg_buffer ----
 #' Function for calculating if the percentage of vegetation within a buffer area is higher than a threshold 
@@
 
-# Function: get_large_polygons ----
+#' Function: get_large_polygons ----
 #' Function for removing small areas from a raster file. Used for calculating large forest areas (e.g., > 5 km2).  
@@
 
-# get_dist_max ----
+#' get_dist_max ----
 #' Get cells within a distance to vegetated area (YES --> 1; NO --> 0)  
@@
 
-# Function: get_wui_raster ----
+#' Function: get_wui_raster ----
 #' Function for WUI classification
@@
 
-# add_colours_wui 
+#' add_colours_wui 
 #' Function for adding WUI colours in a SpatRaster 
@@
 
-Then, we classified as Non-burnable the following categories:
+Then, we classified as non-burnable the following categories:
 
@@
 
-The original resolution of the map is 10 meters. We aggregate it (by a factor of 10) and transform the coordinate system from `SWEREF99 TM (EPSG:3006)` to `ETRS89-extended / LAEA Europe (EPSG:3035)`.
+The original resolution of the map is 10 meters.
+We aggregated it (by a factor of 10) and transformed the coordinate system from `SWEREF99 TM (EPSG:3006)` to `ETRS89-extended / LAEA Europe (EPSG:3035)`.
 
@@
 
-Finally, we combine both maps in one (@fig-wild-vegetation-national)
+Finally, we combined both maps in one (@fig-wild-vegetation-national)
 
@@
 
-Original data from [@Aragoneses2022].
-Al fuels are considered for WUI.
-The resolution of the original map is 1km and we have re-sampled it to a grid cell of 250x250 grids (same than `region_r`) for the WUI (based in the type that is most presented in the area: `mode`).
+Original data from [@Aragoneses2022] and we considered all fuels.
+The resolution of the original map is 1km , so we have re-sampled it to a grid cell of 250x250 metres (same than `region_r`) based in the type that is most presented in the area; i.e., `mode`.
 
@@
 
-We carry out a point-based approach with grid cells of `r raster_resolution` m x `r raster_resolution` m, and a buffer radio of `r radio_buffer` m. Our workflow is therefore:
+We carried out a point-based approach with grid cells of `r raster_resolution` m x `r raster_resolution` m, and a buffer radio of `r radio_buffer` m. Our workflow was therefore:
 
@@
 
-Therefore, we do the following steps:
+Therefore, we did the following steps:
 
@@
 
-Zoom to counties (@fig-wui-map-counties; @tbl-summary-wui-county):
-
-```{r}
-#| label: fig-wui-map-counties
-#| fig-cap: "WUI maps by county based on different fuel/vegetation maps (resolution of 250 x 250 m)"
-#| fig-subcap: 
-#|  - "Viken"
-#|  - "Rogaland"
-#|  - "Vestland"
-#|  - "Trøndelag"
-#|  - "Troms og Finnmark"
-#|  - "Innlandet"
-#|  - "Oslo"
-#|  - "Vestfold og Telemark"
-#|  - "Agder"
-#|  - "Møre og Romsdal"
-#|  - "Nordland"
-#| layout-nrow: 11
-#| out-width: 100%
-
-# plot_zoom <- function(.x) {
-#   
-#   zoom_area <- counties |> 
-#     filter(NUTS_NAME == {{ .x }})
-#   
-#   wui_r_zoom <- wui_r |> 
-#     crop(zoom_area) |> 
-#     mask(zoom_area)
-#   
-#   # Add colour table 
-#   coltab(wui_r_zoom) <- 
-#     data.frame(id = 0:2, 
-#                col = c("lightgrey", "#F0E442", "#CC79A7"))
-#   
-#   par(mfrow = c(1, 4))
-#   
-#   plot(wui_r_zoom$effis, 
-#        axes = FALSE, 
-#        font.main = 1,
-#        legend = FALSE,
-#        mar = c(0, 0, 0, 0),
-#        main =  names(wui_r_zoom$effis))
-#   # title( {{ .x }}, line = -4, adj = 0)
-#   
-#   plot(wui_r_zoom$corine, 
-#        axes = FALSE, 
-#        font.main = 1,
-#        legend = FALSE,
-#        mar = c(0, 0, 0, 0),
-#        main =  names(wui_r_zoom$corine))
-#   
-#   plot(wui_r_zoom$firEUrisk,
-#        axes = FALSE, 
-#        font.main = 1,
-#        legend = FALSE,
-#        mar = c(0, 0, 0, 0), 
-#        main =  names(wui_r_zoom$firEUrisk))
-#   
-#   plot(wui_r_zoom$sr16,
-#        axes = F, 
-#        font.main = 1,
-#        legend = FALSE,
-#        mar = c(0, 0, 0, 0), 
-#        main =  names(wui_r_zoom$sr16)) 
-#   
-# }
-# 
-# # par(mfrow = c(1, 11))
-# as.list(counties$NUTS_NAME) |> 
-#   walk(plot_zoom)
-
-```
+Valuer by counties: @tbl-summary-wui-county):
 
@@
 ```
+
 ```{r}
@@
 
-![](figures/counties_map.png)
-![WUI percentage by county](figures/counties_wui_perc.png)
-
-Area of wild vegetation by county (@tbl-summary-table-areas): Number vegetation cells (centroid is inside the polygon), sum then and calculate the area based on the raster resolution.
+![](figures/counties_map.png) ![WUI percentage by county](figures/counties_wui_perc.png)
+
+Area of wild vegetation by county (@tbl-summary-table-areas): we calculate the number vegetation cells (i.e., grid centroid is inside the polygon), and calculate the area based on the raster resolution.
 
@@
 
-Equivalent to select any area with building as WUI.
+Equivalent to select any area with building as potential WUI.
 
@@
 
-We analyse all wildfires that have affected buildings from 2016 to 2022 (i.e., 330). The data have been obtained from XX (Norway) and XXX (Sweden).
-
 ```{r}
@@
   janitor::clean_names() |> 
-  mutate(datetime = lubridate::as_date(datetime)) |> 
-  filter(datetime >= lubridate::as_date("2016-01-01")) |> 
-  st_as_sf(coords = c("ecoord", "ncoord"),
+  filter(datetime >= 2016) |> 
+  st_as_sf(coords = c("ecoord", "nccord"),
            crs =  3006) |> 
@@
 
-Intersect points with WUI maps
+We have analysed all wildfires that have affected buildings from 2016 to 2023 (i.e., `r dim(wf)[1]`; `r dim(filter(wf, country == "NO"))[1]` in Norway and \``r dim(filter(wf, country == "SE"))[1]` in Sweden).
+The can be obtained from XX (Norway) and XXX (Sweden).
 
@@
 
-Get if the grid to which the wildfire is located was defined as WUI in the respective maps.
+Get if the grid to which the wildfire is located was defined as WUI.
 

### 2024-09-11 article_codes.qmd
Add table with number of wilfires by building density breaks

@@
 
-## Effect distance to large forest
+## Wilfires and building density
+
+Calculate the building density of the grid of 250 x 250 m where the wilfire originated.
+
+```{r}
+#| label: fig-wilfires-number-buildings
+#| fig-cap: Number of buildings per grid cells of 250 m x 250 m and wildfires (red points)
+
+ggplot() +
+  tidyterra::geom_spatvector(data = region_v, fill = "white") +
+  tidyterra::geom_spatraster(data = build_focal) +
+  scale_fill_whitebox_c(name = "N. Buildings",
+                        palette = "viridi",
+                        trans = "log10") +
+  geom_sf(data = wf, col = "red", size = 1) +
+  theme_void()
+
+```
+
+```{r}
+#| label: fig-wildfires-buildings-density
+#| fig-cap: Histogram of building density in each wildfire
+
+# Link wildfires with building density (build_focal) 
+wf_build_density <- extract(build_focal, wf) |>
+  # Assign 0 to NAs --> No buildings
+  mutate(focal_sum = ifelse(is.na(focal_sum), 0, focal_sum)) |> 
+  # Building density is the number of buildings / buffer area 
+  # buffer area = circular moving window of 500 m
+  mutate(bd = focal_sum / area_buffer)
+
+ggplot(wf_build_density, aes(x = bd)) + 
+  geom_histogram(bins = 50, color = "darkblue", fill = "lightblue") + 
+  theme_bw()
+```
+
+```{r}
+#| label: tbl-wildfires-buildings-density
+#| tbl-cap: Number of wildfires
+
+brks <-  c(0,  1, 3, 6, 10, 20, 50, 100, 500, 1700)
+wf_build_density |> 
+  mutate(bd_brks = cut(bd, 
+                       brks,
+                       include.lowest = TRUE,
+                       right = FALSE,
+                       dig.lab = 4)) |> 
+  group_by(bd_brks) |> 
+  summarise(n = n(), 
+            perc = n() / nrow(wf_build_density) * 100) |> 
+  mutate(cs = cumsum(perc)) |> 
+  gt() |> 
+  fmt_number(columns = c("perc", "cs"),
+             decimals = 1) |> 
+  tab_spanner(label = md("**Wildfires**"),
+              columns = c(n, perc, cs)) |> 
+  cols_label(
+    bd_brks = md(html("**Building**<br>**density**")),
+    n  = md("**Number**"),
+    perc = md("**Percentage [%]**"),
+    cs = md("**Cumulative [%]**")
+  )
+
+```
+
+## WUI definition
+
+### Effect distance to large forest
 
@@
 
-## Effect threshold building density
+### Effect threshold building density
 

### 2024-09-18 article_codes.qmd
add touche = TRUE when extracting values from a SpatRaster.
Add a table with number of buildings ein each WUI classification (standard WUI)

@@
 # Summary by county
-tbl_wui_county <- terra::extract(wui_r, counties) |>  
+tbl_wui_county <- terra::extract(wui_r, counties, touches = TRUE) |>  
   pivot_longer(cols = -ID, 
@@
           na.rm = TRUE,
+          touches = TRUE,
           bind = TRUE) |> 
@@
             na.rm = TRUE,
+            touches = TRUE,
             bind = TRUE) |>  
@@
     )
+  )
+
+```
+
+Number of buildings in WUI
+
+```{r}
+#| label: tbl-buildings-wui
+#| tbl-cap: Number of buildings in each WUI classification
+
+c(build_counts, wui_r) |> 
+  as_tibble() |>   
+  mutate(sum = ifelse(is.na(sum), 0, sum)) |> 
+  drop_na() |> 
+  pivot_longer(col = -sum, names_to = "name") |> 
+  group_by(name, value) |>
+  summarise(buildings = sum(sum)) |> 
+  ungroup() |> 
+  pivot_wider(names_from = name, 
+              values_from = buildings) |> 
+  gt(rowname_col =  "value") |> 
+  grand_summary_rows(
+    fns = list(Total = ~sum(.))
+  ) |> 
+   cols_label(
+    corine = md("**corine**"),
+    effis = md("**effis**"),
+    firEUrisk ~ md("**firEUrisk**"),
+    national ~ md("**national**")
   )
@@
 # Link wildfires with building density (build_focal) 
-wf_build_density <- extract(build_focal, wf) |>
+wf_build_density <- extract(build_focal, wf, touches = TRUE) |>
   # Assign 0 to NAs --> No buildings

### 2024-09-20 article_codes.qmd
Add buildigns per WUI (and rerun after cleaning cache!)

@@
   # Add colour table 
-  df_col <- data.frame(id = 0:2, col = c("lightgrey", "#FF6633", "#003366"))
+  df_col <- data.frame(id = 0:2, col = c("#FAFAFA", "#FF6633", "#003366"))
   df_col_lst <- replicate(nlyr(r), df_col, simplify = FALSE)

### 2024-09-28 article.qmd
Remove duplicate coordinates in wildfire dataset

@@
 
-This paper aims to delineate the wildland-urban interface in the sparsely populated, limited resourced Atlantic-Boreal Europe.
-Due to its demographic and biophysical differences to southern Europe, as well as to other fire-prone regions, current WUI definitions to suit the needs of southern countries are not viable.
-Therefore, common WUI mapping assumptions, and how well the resulting WUI maps (thresholds) capture the reality of wildfire in northern Europe, are scoped.
+This paper aims to delineate the wildland-urban interface in the sparsely populated, limited-resourced Scandinavian peninsula (excl.
+Finland).
+Common WUI mapping assumptions and how different thresholds capture the reality of building ignition from wildfire in this region are evaluated.
+We show that dedicated fuel maps capture areas at risk slightly better than vegetation maps, although the choice of map per se is less important than the selected limit for possible ember transport.
+The commonly used 6.17 buildings/km^2^ threshold for building density fails to capture most wildfire incidents that have led to building ignition.
 
@@
 
-This work has been financed by XXX
+N.
+Fernandez-Anez acknowledges The Research Council of Norway, Project 301569: Building Design for At-risk Groups (BUILDER) for supporting her contribution.
+F.
+Vermina Plathner was supported by FORMAS \[grant number 2021-02396\].
+E.
+Pastor acknowledges the Spanish Ministry of Science, Innovation and Universities, Project PID2023-150607OB-I00.

### 2024-09-28 article_codes.qmd
Remove duplicate coordinates in wildfire dataset

@@
 
-Area of wild vegetation by county (@tbl-summary-table-areas): we calculate the number vegetation cells (i.e., grid centroid is inside the polygon), and calculate the area based on the raster resolution.
+Area of wild vegetation by county (@tbl-summary-table-areas): we calculate the number of vegetation cells (i.e., grid centroid is inside the polygon), and calculate the area based on the raster resolution.
 
@@
 
+Total area of WUI (by type of vegetation):
+
+```{r}
+#| label: area-wui-type
+
+wui_r |>  
+  as_tibble() |> 
+  drop_na() |> 
+  pivot_longer(col = everything(),
+               names_to = "name") |> 
+  group_by(name, value) |>
+  summarise(n = n()) |> 
+  ungroup() |> 
+  mutate(area = n * (0.25 * 0.25)) |> #km2
+  pivot_wider(names_from = name, 
+              values_from = c(n, area)) |> 
+  select(value, starts_with("area")) |> 
+  gt() |>
+  fmt_number(decimals = 0) |> 
+  cols_label(
+    value = "",
+    area_corine = md("**corine**"),
+    area_effis = md("**effis**"),
+    area_firEUrisk ~ md("**firEUrisk**"),
+    area_national ~ md("**national**")
+  ) |> 
+  tab_options(
+    table.font.size = 9,
+    row_group.background.color = "lightgrey",
+    row_group.padding = px(1),
+    row_group.font.weight = "bold",
+    data_row.padding = px(1)
+  )
+
+```
+
 Number of buildings in WUI
-
 ```{r}
@@
     national ~ md("**national**")
+  ) |>
+  tab_options(
+    table.font.size = 9,
+    row_group.padding = px(1),
+    row_group.font.weight = "bold",
+    data_row.padding = px(1)
   )
@@
 
-wf_se <- readxl::read_xlsx("data/wui_fires_sweden_epsg3006.xlsx") |> 
+# Infrastructures affected by wildfires in SE
+wf_se_inf <- readxl::read_xlsx("data/wui_fires_sweden_epsg3006.xlsx") |> 
   janitor::clean_names() |> 
@@
 
+# Wildfires in SE (unique coordinates - one wildfire can affect several buildings)
+wf_se <- distinct(wf_se_inf)
+
+# Final dataset (NO + SE)
 wf <- rbind(wf_no, wf_se)
 
+# Map
 ggplot() +
-  geom_sf(data = region_sf, fill = "lightgrey") +
+  geom_sf(data = region_sf, fill = "#FAFAFA") +
   geom_sf(data = country_sf, fill = NA) +
@@
 
-We have analysed all wildfires that have affected buildings from 2016 to 2023 (i.e., `r dim(wf)[1]`; `r dim(filter(wf, country == "NO"))[1]` in Norway and \``r dim(filter(wf, country == "SE"))[1]` in Sweden).
-The can be obtained from XX (Norway) and XXX (Sweden).
+We have analysed all wildfires that have affected buildings from 2016 to 2023 (i.e., `r dim(wf)[1]`; `r dim(filter(wf, country == "NO"))[1]` in Norway and `r dim(filter(wf, country == "SE"))[1]` in Sweden).
+The can be obtained from the *Norwegian Directorate for Civil Protection (DSB)*  (Norway) and the *Swedish Civil Contingencies Agency (MSB)* (Sweden).
+[Note: They were not uploaded in our GitHub repository]
 
@@
     cs = md("**Cumulative [%]**")
+  ) |> 
+  tab_options(
+    table.font.size = 9,
+    row_group.padding = px(1),
+    row_group.font.weight = "bold",
+    data_row.padding = px(1)
   )

### 2024-09-30 article_codes.qmd
Correct typos and links to figures/tables

@@
 The resolution of the raster is 100 m x 100 m.
-We classify as potential wildland vegetation the following classes @fig-corine-wild-vegetation:
+We classify as potential wildland vegetation the following classes @fig-corine-fuel-map-norway:
 
@@
 
-Original data from [@Aragoneses2022] and we considered all fuels.
-The resolution of the original map is 1km , so we have re-sampled it to a grid cell of 250x250 metres (same than `region_r`) based in the type that is most presented in the area; i.e., `mode`.
+We use the fuel map produced by [@Aragoneses2022].
+The original resolution of the map is 1km, so we resampled it to a grid cell of 250x250 metres (same grids than `region_r`) based in the type that is most presented in the area; i.e., `mode` (@fig-firEUrisk-fuel-map-norway).
 
@@
 
-Valuer by counties: @tbl-summary-wui-county):
+Values by counties:
 

### 2024-11-26 article.qmd
Update Acknowledgements

@@
 
-N.
+This work was partially supported by an Short-Term Scientific Mission Grant from the COST Action CA18134 FIRElinks.N.
 Fernandez-Anez acknowledges The Research Council of Norway, Project 301569: Building Design for At-risk Groups (BUILDER) for supporting her contribution.
@@
 Pastor acknowledges the Spanish Ministry of Science, Innovation and Universities, Project PID2023-150607OB-I00.
+Building data have been downloaded from Geofabrik GmbH (<https://www.geofabrik.de/>).
+Map data copyrighted OpenStreetMap contributors and available from [https://www.openstreetmap.org](https://www.openstreetmap.org/).
+This publication has been prepared using European Union's Copernicus Land Monitoring Service information: <https://doi.org/10.2909/960998c1-1870-4e82-8051-6485205ebbac>.
